---
title: "Regional Housing Framework"
subtitle: "Clean Arlington County public records data"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

#### Library: RegHsg
#### Project: Regional Housing Framework
#### Author: Sarah Strochak
#### Version: R 3.5.1, RStudio 1.1.423
#### Last updated `r format(Sys.time(), '%B %d, %Y')`
#### Environment: Local Windows session (desktop)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

##### There are three main goals of the jurisdiction level cleaning process:
1. Recategorize the county land use codes into more general codes
2. Collapse observations at the address level
3. Clean the variables needed to provide density estimates

## Set-up
### Load libraries and functions
```{r setup}
library(tidyverse)
library(DescTools)
library(purrr)

source("../Macros/read-bk.R")
source("../Macros/filter-bk.R")
source("../Macros/select-vars.R")
source("../Macros/sample-properties.R")
source("../Macros/collapse-properties.R")
```

#### Set FIPS code and filepath name
```{r fips}
currentfips <- "51013"
filepath <- "arlington"
```

#### Load in Black Knight data for the region, select jurisdiction and standard variables
```{r read}
if (!exists("region")) {
  region <- read_bk_rmd("dc-cog-assessment_20181228.csv")
} else {
  warning("region data already read in")
}

currentjur <- region %>% 
  filter_bk(fips = currentfips) %>% 
  select_vars()

```
## Recatagorize county land use codes
##### In order to complete the vacant land and soft site analysis, we will need to break down properties into different classifications. This process creates three variables:
1. residential: 1 for residential, 0 for other
2. category: categories are sf, mf, commercial, office, and vacant.
3. category_detail: this will vary by jurisdiction, but  includes the most detail possible to generalize from the county land use codes.

#### Export county land use codes for manual classification
```{r export-landuse}
currentjur_county <- currentjur %>% 
  group_by(countylandusedescription) %>% 
  count()

if (!file.exists(paste0("../Data/", filepath, "-county-land-use.csv"))) {
  write_csv(currentjur_county,
            paste0("../Data/", filepath, "-county-land-use.csv"))
}

```

##### Create variables using `ifelse()` and `case_when()`
```{r classify}
res_codes <- 
  c("AFFORDABLE DWELLING UNIT",
    "APARTMENT - GARDEN",
    "APARTMENT - HIGH-RISE",
    "APARTMENT - MID-RISE",
    "APARTMENT - PARKING",
    "COMMUNITY BENEFIT UNIT",
    "MULTI-FAM IMPR/NO SITE PLAN",
    "MULTI-FAM IMPR/SITE PLAN",
    "MULTI-FAM VACANT/NO SITE PLAN",
    "MULTI-FAM VACANT/SITE PLAN",
    "NOT VALUED CONDO HOA",
    "NOT VALUED COSTED - HOA/SITE PLAN",
    "RESIDENTIAL COST-VAL - DUPLEX",
    "RESIDENTIAL COST-VAL - IMPR/SF & TW",
    "RESIDENTIAL COST-VAL - SIDE BY SIDE",
    "RESIDENTIAL COST-VAL - SINGLE-FAM DET",
    "RESIDENTIAL COST-VAL - TOWNHOUSE/CO",
    "RESIDENTIAL COST-VAL - TOWNHOUSE/FSO",
    "RESIDENTIAL COST-VAL - VAC/SF & TW",
    "SALES APPR CONDO/CO-OP",
    "SALES APPR CONDO/GARDEN",
    "SALES APPR CONDO/HIGH RISE",
    "SALES APPR CONDO/MID RISE",
    "SALES APPR CONDO/STACKED",
    "SFD - APT ZONED/NO SITE PLAN",
    "SFD - APT ZONED/SITE PLAN",
    "VACANT RESIDENTIAL")

currentjur_cat <- currentjur %>% 
  mutate(residential =
           ifelse(countylandusedescription %in% res_codes, 1, 0),
         category = case_when(
           countylandusedescription %in%
             c("COMMERCIAL CONDO",
               "GEN COMM-IND - AUTO DEALERSHIP",
               "GEN COMM-IND - SELF-STORAGE",
               "GEN COMM-IND - SERVICE STATION",
               "GEN COMM-IND - WAREHOUSE",
               "GEN COMM - BANK",
               "GEN COMM - FAST FOOD",
               "GEN COMM - HEALTH CARE FACILITY",
               "GEN COMM - MIXED OFFICE/COMM",
               "GEN COMM - NEIGHBORHOOD CTR",
               "GEN COMM - OTHER",
               "GEN COMM - REST/EATING FACILITY",
               "GEN COMM - RETAIL/STRIP",
               "GEN COMM - SMALL OFFICE",
               "GEN COMM IMPR-LAND/NO SITE PLAN",
               "GEN COMM IMPR-LAND/SITE PLAN",
               "GEN COMM/PARKING",
               "HOTEL - FULL SERVICE",
               "HOTEL - LAND/OTHER",
               "HOTEL - LIMITED SERVICE",
               "HOTEL - LODGING",
               "HOTEL - SELECT DRIVE",
               "HOTEL RESIDENCE SUITES",
               "MIXED USE",
               "SFD - COMM ZONED/NO SITE PLAN",
               "SFD - COMM ZONED/SITE PLAN") ~ "commercial",
           countylandusedescription %in%
             c("AFFORDABLE DWELLING UNIT",
               "APARTMENT - GARDEN",
               "APARTMENT - HIGH-RISE",
               "APARTMENT - MID-RISE",
               "APARTMENT - PARKING",
               "COMMUNITY BENEFIT UNIT",
               "MULTI-FAM IMPR/NO SITE PLAN",
               "MULTI-FAM IMPR/SITE PLAN",
               "NOT VALUED CONDO HOA",
               "RESIDENTIAL COST-VAL - DUPLEX",
               "SALES APPR CONDO/CO-OP",
               "SALES APPR CONDO/GARDEN",
               "SALES APPR CONDO/HIGH RISE",
               "SALES APPR CONDO/MID RISE",
               "SALES APPR CONDO/STACKED",
               "SFD - APT ZONED/NO SITE PLAN",
               "SFD - APT ZONED/SITE PLAN") ~ "mf",
           countylandusedescription %in%
             c("OFFICE BLDG IMPR-LAND/SITE PLAN",
               "OFFICE BLDG/7 OR MORE STORIES",
               "OFFICE BLDG/PARKING",
               "OFFICE BLDG/UNDER 7 STORIES") ~ "office",
           countylandusedescription %in%
             c("NOT VALUED COSTED - HOA/SITE PLAN",
               "RESIDENTIAL COST-VAL - IMPR/SF & TW",
               "RESIDENTIAL COST-VAL - SIDE BY SIDE",
               "RESIDENTIAL COST-VAL - SINGLE-FAM DET",
               "RESIDENTIAL COST-VAL - TOWNHOUSE/CO",
               "RESIDENTIAL COST-VAL - TOWNHOUSE/FSO") ~ "sf",
           countylandusedescription %in% 
             c("MULTI-FAM VACANT/NO SITE PLAN",
               "MULTI-FAM VACANT/SITE PLAN",
               "RESIDENTIAL COST-VAL - VAC/SF & TW",
               "VACANT RESIDENTIAL",
               "GEN COMM VAC-LAND/NO SITE PLAN",
               "GEN COMM VAC-LAND/SITE PLAN",
               "OFFICE BLDG VAC-LAND/NO SITE PLAN",
               "OFFICE BLDG VAC-LAND/SITE PLAN",
               "VACANT COMMERCIAL",
               "VACANT EXEMPT",
               "VACANT LIGHT INDUSTRIAL",
               "VACANT OFFICE") ~ "vacant"),
         category_detail = case_when(
           countylandusedescription %in%
             c("AFFORDABLE DWELLING UNIT",
               "APARTMENT - GARDEN",
               "APARTMENT - HIGH-RISE",
               "APARTMENT - MID-RISE",
               "APARTMENT - PARKING",
               "COMMUNITY BENEFIT UNIT",
               "MULTI-FAM IMPR/NO SITE PLAN",
               "MULTI-FAM IMPR/SITE PLAN",
               "SFD - APT ZONED/NO SITE PLAN",
               "SFD - APT ZONED/SITE PLAN") ~ "apartment",
           countylandusedescription %in% 
             c("COMMERCIAL CONDO",
               "GEN COMM-IND - AUTO DEALERSHIP",
               "GEN COMM-IND - SELF-STORAGE",
               "GEN COMM-IND - SERVICE STATION",
               "GEN COMM-IND - WAREHOUSE",
               "GEN COMM - BANK",
               "GEN COMM - FAST FOOD",
               "GEN COMM - HEALTH CARE FACILITY",
               "GEN COMM - MIXED OFFICE/COMM",
               "GEN COMM - NEIGHBORHOOD CTR",
               "GEN COMM - OTHER",
               "GEN COMM - REST/EATING FACILITY",
               "GEN COMM - RETAIL/STRIP",
               "GEN COMM - SMALL OFFICE",
               "GEN COMM IMPR-LAND/NO SITE PLAN",
               "GEN COMM IMPR-LAND/SITE PLAN",
               "GEN COMM/PARKING",
               "HOTEL - FULL SERVICE",
               "HOTEL - LAND/OTHER",
               "HOTEL - LIMITED SERVICE",
               "HOTEL - LODGING",
               "HOTEL - SELECT DRIVE",
               "HOTEL RESIDENCE SUITES",
               "MIXED USE",
               "SFD - COMM ZONED/NO SITE PLAN",
               "SFD - COMM ZONED/SITE PLAN") ~ "commercial",
           countylandusedescription %in% 
             c("NOT VALUED CONDO HOA",
               "SALES APPR CONDO/CO-OP",
               "SALES APPR CONDO/GARDEN",
               "SALES APPR CONDO/HIGH RISE",
               "SALES APPR CONDO/MID RISE",
               "SALES APPR CONDO/STACKED") ~ "condo",
           countylandusedescription %in% 
             c("RESIDENTIAL COST-VAL - DUPLEX") ~ "duplex",
           countylandusedescription %in%
             c("OFFICE BLDG IMPR-LAND/SITE PLAN",
               "OFFICE BLDG/7 OR MORE STORIES",
               "OFFICE BLDG/PARKING",
               "OFFICE BLDG/UNDER 7 STORIES") ~ "office",
           countylandusedescription %in%
             c("RESIDENTIAL COST-VAL - SIDE BY SIDE") ~ "sf attached",
           countylandusedescription %in% 
             c("NOT VALUED COSTED - HOA/SITE PLAN",
               "RESIDENTIAL COST-VAL - IMPR/SF & TW",
               "RESIDENTIAL COST-VAL - SINGLE-FAM DET") ~ "sf detached",
           countylandusedescription %in% 
             c("RESIDENTIAL COST-VAL - TOWNHOUSE/CO",
               "RESIDENTIAL COST-VAL - TOWNHOUSE/FSO") ~ "townhouse",
           countylandusedescription %in% 
             c("GEN COMM VAC-LAND/NO SITE PLAN",
               "GEN COMM VAC-LAND/SITE PLAN",
               "VACANT COMMERCIAL") ~ "vacant commercial",
           countylandusedescription %in% 
             c("VACANT EXEMPT") ~ "Vacant exempt",
           countylandusedescription %in% 
             c("VACANT LIGHT INDUSTRIAL") ~ "vacant light industrial",
           countylandusedescription %in%
             c("MULTI-FAM VACANT/NO SITE PLAN",
               "MULTI-FAM VACANT/SITE PLAN") ~ "vacant mf",
           countylandusedescription %in%
             c("OFFICE BLDG VAC-LAND/NO SITE PLAN",
               "OFFICE BLDG VAC-LAND/SITE PLAN",
               "VACANT OFFICE") ~ "vacant office",
           countylandusedescription %in%
             c("VACANT RESIDENTIAL") ~ "vacant residential",
           countylandusedescription %in%
             c("RESIDENTIAL COST-VAL - VAC/SF & TW") ~ "vacant sf")
         )
```

#### These checks make sure that all county land use codes recieved new classifications, and shows the distribution of properties 

#### {.tabset .tabset.fade}
##### residential
```{r res-check}
count(currentjur_cat, residential)
count(currentjur_cat, is.na(residential))
```

##### category
```{r cat-check}
count(currentjur_cat, category)
count(currentjur_cat, is.na(category))
```
##### category_detail
```{r catd-check}
count(currentjur_cat, category_detail)
count(currentjur_cat, is.na(category_detail))
```

## Collapsing properties
##### Next, we have to ensure that there is only one observation per address in the dataset. 

##### There are a number of properties that have numbers or symbols in their house numbers that prevent a collapse. This next step removes those characters so the collapse is clean.

```{r letters}
currentjur_cat2 <- currentjur_cat %>% 
  mutate(house_letter = ifelse(str_detect(prophouseno, "[:alpha:]") == 1, 1, 0),
         oldadd = propaddress,
         new_houseno = ifelse(house_letter == 1,
                              str_replace(prophouseno, "[:alpha:]", ""), prophouseno),
         propaddress = ifelse(house_letter == 1, 
                              paste(str_replace_all(new_houseno, "-", ""),
                                    propstreetname, 
                                    propstreetsuffix,
                                    sep = " "),
                              propaddress),
         propaddress = ifelse(is.na(propaddress), oldadd, propaddress))
```

##### Compare the first 20 addresses to check
```{r letters-check}
currentjur_cat2 %>% 
  filter(house_letter == 1) %>% 
  select(oldadd, propaddress) %>% 
  head(20)

```

##### An additional check to make sure no new NAs in the address were introduced
```{r letters-check2}
if (sum(is.na(currentjur_cat2$propaddress)) > 
    sum(is.na(currentjur_cat$propaddress))) {
  warning("additional NAs are introduced") 
  } else if (sum(is.na(currentjur_cat2$propaddress)) < 
             sum(is.na(currentjur_cat$propaddress))) {
  warning("less NAs than original")
    } else if (sum(is.na(currentjur_cat2$propaddress)) == 
               sum(is.na(currentjur_cat$propaddress))) {
      print("no additional NAs")
    }
```

##### In order to do the collapse, we first identify addresses with more than one observation, and filter out those with a missing house number or address to avoid an improper collapse. Then, run a quick check to make sure all the records are still accounted for.

```{r sep}
singles <- get_single_properties(currentjur_cat2)
multiples <- get_multiple_properties(currentjur_cat2)
missing_address <- get_missing_address(currentjur_cat2)
check_classification(currentjur_cat2)
```

##### For addresses that have some records with missing zoning, but others that have a zoning designation: we need to fill the NAs with the appropriate zoning code. `group_by()` means that zoning codes will not be passed on to observations with a different address, and using `fill()` in both directions means that the NAs will be filled regardless of their position.

```{r fill}
multiples <- multiples %>% 
  group_by(propaddress) %>% 
  fill(zoning) %>% 
  fill(zoning, .direction = "up") %>% 
  ungroup()
```

##### Next, we summarize the data using nested lists, which allow us to preserve all the details from each record and apply the right function to each. The table below describes each variable, which function to use, and what to do in the case of a tie. Lot size is a unique case, detail provided in the next section.

```{r var-table, echo=FALSE, results = 'asis', eval=TRUE}

v <- tibble::tibble(variable = c("zoning", 
                         "lotsizeorarea", 
                         "lotsizesquarefeet",
                         "buildingarea", 
                         "countylandusedescription",
                         "residential", 
                         "category", 
                         "category_detail",
                         "yearbuilt", 
                         "long", 
                         "lat"),
            operation = c("mode", 
                          "conditional based on values",
                          "conditional based on values",
                          "sum",
                          "mode",
                          "max",
                          "mode",
                          "mode",
                          "max",
                          "median",
                          "median"),
            `in case of tie` = c("take most dense zoning code (?)",
                                 "",
                                 "",
                                 "",
                                 "case-by-case basis",
                                 "residential",
                                 "case-by-case basis",
                                 "case-by-case basis",
                                 "take maxiumum year- accounts for renovation",
                                 "median longitude and latitude ensures this is a correct pairing, falls within the address",
                                 "median longitude and latitude ensures this is a correct pairing, falls within the address"))

knitr::kable(v)

```


``` {r nest}
nested <- multiples %>% 
  group_by(propaddress) %>%
  summarise_at(vars(zoning, lotsizeorarea, lotsizesquarefeet,
                    buildingarea, countylandusedescription,
                    residential, category, category_detail,
                    yearbuilt, long, lat), list) %>% 
  rename_at(vars(-propaddress), ~ paste0(., "_list")) %>% 
  mutate(nprops = map(zoning_list, length),
         zoning = map(zoning_list, Mode),
         lotsizeorarea = map(lotsizeorarea_list, sum),
         lotsizesquarefeet = map(lotsizesquarefeet_list, sum),
         buildingarea = map(buildingarea_list, sum),
         countylandusedescription = map(countylandusedescription_list, Mode),
         residential = map(residential_list, max),
         category = map(category_list, Mode),
         category_detail = map(category_detail_list, Mode),
         yearbuilt = map(yearbuilt_list, max),
         long = map(long_list, median),
         lat = map(lat_list, median))
```