---
title: "Regional Housing Framework"
subtitle: "Post Clean Montgomery BK data and join it with county data"
description: "takes the collapsed parcel file and spatially joins it with a county parcel file to fill in missing lot size and zoning variables. Merges with rockville and gaithersburg shape files to create analysis flags for use in the vacant lots analysis"
author: "Patrick Spauster"
date: "February 13, 2019"
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Patrick Spauster

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

## Set-up


Load libraries and functions

```{r setup}
library(tidyverse)
library(DescTools)
library(purrr)
library(sf)

source("../Macros/read-bk.R")
source("../Macros/filter-bk.R")
source("../Macros/select-vars.R")
source("../Macros/sample-properties.R")
source("../Macros/classify-addresses.R")
source("../Macros/max_narm.R")
source("../Macros/mode_narm.R")
```

Set FIPS code and filepath name

```{r fips}
currentfips <- "24031"
filepath <- "montgomery"
jdir <- paste0("L:/Libraries/RegHsg/Data/", filepath, "/")
rdir <- paste0("L:/Libraries/RegHsg/Raw/", filepath, "/")
```

Load in cleaned Black Knight data.

```{r readR}
jur <- readRDS(paste0(jdir, 
               "cleaned-",
               filepath,
               "-data.Rdata"))
```



## Supplement with county data

Use county data sources to fill in missing information.

Read in zoning file (zoning_sf) to match to the collapsed BK data.
Read in property file and master address file.

```{r readin}

zoning_sf <- read_sf(dsn = 'L:/Libraries/RegHsg/Raw/montgomery/montgomery-parcel-file/MCzoning_parcels',
                     layer = "MC_zoning")

# property shape stored in geodatabase- assign gdb for extraction
prop <- "L:/Libraries/RegHsg/Maps/montgomery/property/property.gdb"

# function below lists later in geodatabase
# rgdal::ogrListLayers(prop)

propsf <- read_sf(prop, layer = "Property_poly_with_data")

# read in master address file
ma <- read_csv("L:/Libraries/RegHsg/Data/masteraddress_gis.csv",
               col_types = cols(
  ADDRID = col_double(),
  ADDR_STATUS = col_double(),
  ADDR_TYPE = col_double(),
  ADDR_DATE = col_character(),
  ADDR_NUMBER = col_double(),
  ADDR_NUMBER_SUFFIX = col_character(),
  PRE_DIR = col_character(),
  ST_NAME = col_character(),
  ST_TYPE = col_character(),
  SUF_DIR = col_character(),
  UNIT = col_character(),
  XCOORD = col_double(),
  YCOORD = col_double(),
  FULL_ADDRESS = col_character(),
  CITY = col_character(),
  ZIPCODE = col_double(),
  COMMENTS = col_character(),
  COUNTY_CODE = col_double(),
  DISTRICT = col_character(),
  ACCT = col_character(),
  SUBDIVISION = col_character(),
  SUBDIV_ID = col_double(),
  SUBDIV_NAME = col_character(),
  LOT = col_character(),
  BLOCK = col_character(),
  PARCEL_NO = col_character(),
  LU_CATEGORY = col_character(),
  CBPTEAM = col_double(),
  LONGZONE = col_character(),
  MASTERPLAN_CODE = col_character(),
  MASTERPLAN = col_character(),
  POLICY_NUMBER = col_double(),
  POLICY_AREA = col_character(),
  PRELIM_APNO = col_character(),
  SITE_APNO = col_character(),
  SAR_STATUS = col_character(),
  CHAMPTREE = col_character(),
  SFP_ES = col_character(),
  SFP_MS = col_character(),
  SFP_HS = col_character(),
  TPAR_REQ = col_character(),
  PAMR_REQ = col_character(),
  CONDITION_APNOS = col_character(),
  CHANGE_DATE = col_character(),
  X45 = col_character()
))

```

Use the property file to supplement lot size for parcels with no "C" in their ACCT. For those who do not match on this- supplement parcels with a "C" with the address file, which crosswalks addresses to parcels with "C"s. Then re-merge to the property file, and divide the area of the "C" parcels by the number of observations per parcel ID.

```{r}
# get rid of geomerty for processing time
prop_testing <- propsf %>% 
  as.data.frame()

# select and create variables from property file to merge with black knight data
# select only obs with unique parcel ID
propmerge <- prop_testing %>% 
  select(DISTRICT, ACCT, 
         LU_CATEGORY,
         CONDO_UNIT_NO, 
         NO_DWELLINGS, DWELLING_TYPE,
         PLANNING, POLICY, MASTER_PLAN,
         RES_DWELLU, LAND_ASSMT, IMPROV_ASSMT, Shape_Area) %>% 
  mutate(parcel_id = paste0(DISTRICT, "-", ACCT),
         cflag = str_detect(ACCT, "C"),
         propflag = 1) %>% 
  group_by(parcel_id) %>% 
  filter(n() == 1) %>% 
  ungroup()

# join black knight with property file
jur1 <- jur %>% 
  mutate(missing_houseno = ifelse(is.na(prophouseno) &
                                    address_type == "single",
                                  1,
                                  0)) %>% 
  left_join(propmerge, by = c("assessorsparcelnumberapnpin" = "parcel_id"))

# from master address file- select obs that are distict address + parcel ID pairs
mac <- ma %>% 
  filter(str_detect(ACCT, "C")) %>% 
  group_by(FULL_ADDRESS) %>% 
  summarize(acct_list = list(ACCT)) %>% 
  mutate(num = map_dbl(acct_list, ~ length(unique(.))),
         house1 = substr(FULL_ADDRESS, 1, 2)) %>% 
  filter(num == 1,
         # get rid of observations with house numbers of 0
         !house1 == "0 ") %>% 
  mutate(ACCT_address = map_chr(acct_list, 1))

# use this as parcel-address crosswalk
macmerge <- mac %>% 
  select(FULL_ADDRESS, ACCT_address)

# join only obs that did not match in parcel file and do not have missing addresses
jur1a <- jur1 %>% 
  filter(is.na(propflag),
         missing_houseno == 0) %>% 
  select(-c(38:48)) %>% 
  left_join(macmerge, by = c("propaddress" = "FULL_ADDRESS"))

# see how many obs in the black knight data did not have a match in the master address file
jur1 %>% 
  filter(is.na(propflag),
         missing_houseno == 0) %>% 
  anti_join(mac, by = c("propaddress" = "FULL_ADDRESS")) %>% 
  nrow()

# join by parcel ID back to property file
jur1b <- jur1a %>% 
  left_join(propmerge, by = c("ACCT_address" = "ACCT")) %>% 
  group_by(ACCT_address) %>% 
  # divide area by the number of parcels
  mutate(lot_acct = Shape_Area.y / n(),
         clotflag = 1) %>% 
  ungroup()

# test how many and which type of properties did not join
jur1banti <- jur1a %>% 
  anti_join(propmerge, by = c("ACCT_address" = "ACCT")) %>% 
  count(category_detail)

```
## Fill lot size.

Recombine and replace lot size with the appropriate measure.

```{r recombine}

# recombine with obs that joined to property file or have missing house number
jur2 <- jur1 %>% 
  filter(propflag == 1 & !is.na(propflag) |
         missing_houseno == 1) %>% 
  bind_rows(jur1b) %>% 
  # create a flag for when the subsetted area should be used
  mutate(clotflag = ifelse(is.na(clotflag),
                           0,
                           clotflag))


jur3 <- jur2 %>% 
  mutate(arc_lotsize = lotsize_sf,
         lotsize_sf = case_when(
    # if subsetted area is available, use this
    clotflag == 1 ~ lot_acct,
    # if area was not subsetted and original lot size is missing, use the shape area
    clotflag == 0 & is.na(arc_lotsize) ~ Shape_Area,
    # if neither of these things are true, keep black knight lot size
    TRUE ~ lotsize_sf
  ))

# test how many properties have missing lot size
jur3 %>% 
  group_by(category_detail) %>% 
  summarize(n(),
            sum(is.na(lotsize_sf)),
            sum(lotsize_sf == 0, na.rm = TRUE)) %>% 
  knitr::kable()
```

Look at the difference between shape area and the original black knight lot size- make sure it is not severly skewed one way.

```{r}
jur3 %>% 
  mutate(lsdiff = Shape_Area - arc_lotsize) %>% 
  filter(!is.na(arc_lotsize),
         !is.na(Shape_Area),
         clotflag != 1) %>% 
  ggplot() +
  geom_histogram(mapping = aes(lsdiff)) +
  scale_x_continuous(limits = c(-500, 500))

```

### Fix land use

```{r fix-landuse}

landuse <- pull(prop_testing, LU_CATEGORY) %>% unique() %>% sort()

landuse_recode <- tibble(LU_CATEGORY = landuse,
                         landuse_residential = 
                           c(0, 1, 0, NA, 0, 0, 
                             1, 0, 0, 0, 0, 0,
                             0, 1, 1, 0, 0, 0),
                         landuse_category = c("other", "mf", "other",
                                              NA, "other", "other",
                                              "mf", "other", "other",
                                              "other", "other", "other",
                                              "other", "sf", "sf",
                                              "other", "vacant", "other"))

jur3a <- jur3 %>% 
  left_join(landuse_recode, by = "LU_CATEGORY")

jur3a %>% 
  count(residential == landuse_residential)

jur3a %>% 
  count(category == landuse_category)

jur3a <- jur3a %>% 
  mutate(residential = ifelse(!is.na(landuse_residential) &
                                residential != landuse_residential,
                              landuse_residential,
                              residential),
         category = ifelse(!is.na(landuse_category) &
                           category != landuse_category,
                           landuse_category,
                           category))

jur3a %>% 
  count(residential == landuse_residential)

jur3a %>% 
  count(category == landuse_category)

jur3a %>% 
  count(category)

```


## Spatial joins

Next, use a spatial join to make flags for Rockville and Gaithersburg.

Convert Black Knight data to spatial.

```{r cleanjur}
jur4 <- jur3a %>% 
  mutate(missing_coord = ifelse(is.na(lat) | is.na(long),
                                1,
                                0))

jur4 <- jur4 %>% #archive the lat and long so that we can do multiples spatial joins. SF replace lat long with an SF object
  mutate(latarc = lat) %>% 
  mutate(longarc = long) %>% 
  mutate(lat1 = lat) %>% 
  mutate(long1 = long) %>% 
  mutate(lat2 = lat) %>% 
  mutate(long2 = long)


jur_sf <- jur4 %>% #set geometry
  filter(missing_coord == 0) %>% 
  st_as_sf(coords = c("long", "lat")) %>% 
  st_set_crs(st_crs(4326))

```


### Zoning code: fill zoning

Find the invalid geometries in the zoning code

```{r find bad geometries zoning}

zoning_sf1 <- zoning_sf %>% 
  # there is one large shape overlayed over all the other shapes- filter this out
  filter(PERIMETER == 0) %>%
  mutate(validgeo = ifelse(st_is_valid(st_sfc(geometry)),
                                1,
                                0))

zoning_sf1 %>% as.data.frame() %>% 
  count(validgeo)

# transform CRS to be consistent with Black Knight data
zoning_sf1 <- st_transform(zoning_sf1, crs = 4326)

```


Join the Black Knight data to the zoning code to fill in missing instances.

```{r join zoning onto parcel}

zjoin <- st_join(jur_sf, zoning_sf1, join = st_intersects)

# Make sure test variable is never missing
zoning_sf1 %>% as.data.frame() %>% 
  count(is.na(ZONE_))

zjoin %>% 
  as.data.frame() %>% 
  count(is.na(ZONE_))

zjoin <- zjoin %>% 
  mutate(notinmont = if_else(is.na(ZONE_), 1, 0))

zjoin %>%  
  as.data.frame() %>% 
  count(notinmont)


```


### Jurisdiction flags

Our Data on the property city is not reliable, so we need to create a flag ourselves using the Rockville and Gaithersburg zoning layers to.

#### Rockville

I read in a rockville zoning file, and add a flag for any property that falls within the Rockville shape file. This will help us filter and use the right zoning code in the vacant lots analysis

Read file, test geometry and projection.

```{r read tiger line 2018 place shapefile}

#read in tiger line place file to create flags for places with their own zoning codes

placepfile <- str_sub(list.files('L:/Libraries/RegHsg/Raw/montgomery/placefile'), 
                     end = -5) %>% unique()
Mocoplaces_sf <- read_sf(dsn = 'L:/Libraries/RegHsg/Raw/montgomery/placefile',
                     layer = placepfile[1])

Mocoplaces_sf1 <- Mocoplaces_sf %>% 
  mutate(validgeo = ifelse(st_is_valid(st_sfc(geometry)),
                                1,
                                0))
Mocoplaces_sf1 %>% as.data.frame() %>% 
  count(validgeo)

st_geometry(Mocoplaces_sf1)

Mocoplaces_sf2 <- st_transform(Mocoplaces_sf1, crs = 4326)


pjoin <- st_join(zjoin, Mocoplaces_sf2, join = st_intersects)

pjoin2 <- pjoin %>% 
  # make Rockville flag
  mutate(rockflag = if_else(NAME=="Rockville",1,0),
         rockflag = if_else(is.na(NAME),0,rockflag),
         gaitflag = if_else(NAME=="Gaithersburg",1,0),
         gaitflag = if_else(is.na(NAME),0,gaitflag),
         othernotmoco = if_else(NAME %in% c("Barnesville","Brookville","Laytonsville","Poolesville"),1,0),
         othernotmoco = if_else(is.na(NAME),0,othernotmoco)
        
  )
count(pjoin2,rockflag)
count(pjoin2,gaitflag)
count(pjoin2,othernotmoco)

```

```{r test-rockville-zoning}

rockpfile <- str_sub(list.files('L:/Libraries/RegHsg/Raw/montgomery/Rockville_zoning_districts'), 
                     end = -5) %>% unique()

rockzoning_sf <- read_sf(dsn = 'L:/Libraries/RegHsg/Raw/montgomery/Rockville_zoning_districts',
                     layer = rockpfile[1])

rockzoning_sf1 <- rockzoning_sf %>% 
  mutate(validgeo = ifelse(st_is_valid(st_sfc(geometry)),
                                1,
                                0))
rockzoning_sf1 %>% as.data.frame() %>% 
  count(validgeo)
rockzoning_sf1 <- rockzoning_sf1 %>% 
  filter(validgeo == 1 & !is.na(validgeo))

st_geometry(rockzoning_sf1)
```

```{r join rockville}

# remove overlap in the zoning shapefile
rzoning <- st_intersection(rockzoning_sf1) %>% 
  filter(n.overlaps == 1)


rjoin <- st_join(pjoin2, rzoning, join = st_intersects)

```

```{r test join rockville}

paste0(sum(!is.na(rjoin$ZONE)), " observations joined- ",
      round(sum(!is.na(rjoin$ZONE)) / nrow(rjoin) * 100, 2), "%")

```

16150 observations joined, so we estimate there are that many parcels in Rockville.

##### Replace BK Zoning with Rockvile Zoning 

```{r filter to make flag}

rjoin2 <- rjoin %>% 
  # replace moco zoning with rockville zoning
  mutate(zoning = if_else(rockflag == 1, ZONE, zoning))

rjoin2 %>% 
  as.data.frame() %>% 
  filter(rockflag ==1) %>% 
  count(zoning)

# test to see that these parcels are actually in rockville =
test <- rjoin2 %>% 
  filter(rockflag ==1) %>% 
  sample_n(1000)

ggplot() +
  geom_sf(rockzoning_sf1, mapping = aes()) +
  geom_sf(test, mapping = aes())


# see how this compares to result of moco zoning join
rjoin2 %>%  
  as.data.frame() %>% 
  count(notinmont, rockflag)

```


After the join, we convert back to a dataframe, for ease of computations.
```{r back-to-df again}

jur5 <- st_set_geometry(rjoin2, NULL)

```


There are 4046 observations that we think are not in montgomery codes, rockville, or gaithersburg.

```{r}
jur5 %>% 
  filter(notinmont == 1 &
           rockflag == 0 &
           gaitflag == 0) %>% 
  nrow()

```


## Missing zoning variables

To determine zoning codes for addresses that have a missing zoning designation or multiple zoning designations, we will use a spatial join to the county's zoning layer.


### Missing zoning

```{r fill-lots zoning}

jur6 <- jur5 %>% 
  # archive the original zoning variable
  mutate(arc_zoning = zoning, 
         # fill in whatever we can from the montgomery shape file if it's missing
         zoning = ifelse(is.na(zoning),
                             CODE,
                             zoning)
         
         # we already replaced this so we don't want to use it to fill in right?
         #fill in whatever we can from the rockville shape file if it's missing
         #zoning = ifelse(is.na(zoning),
                             #ZONE,
                            #  zoning)
         )
        

```

See how many were filled in

```{r count-lotsZ}
sum(is.na(jur6$arc_zoning)) - sum(is.na(jur6$zoning))

# look at remaining missing zoning
jur6 %>% 
  filter(is.na(zoning) & othernotmoco==0) %>% 
  count(category_detail)

```



Test how many have different results from the Black Knight zoning variable.

```{r test-spatialZ}

jur6 %>% 
  group_by(is.na(zoning)) %>% 
  count(zoning == CODE)

```



```{r missingZ}

jur6 %>% 
  mutate(num = NA) %>% 
  group_by(category_detail) %>% 
  summarize_at(vars(propaddress, lotsize_sf, zoning, num),
               ~ sum(is.na(.))) %>% 
  mutate_at(vars(propaddress, lotsize_sf, zoning),
            .funs = funs(percent = scales::percent(. / num))) %>% 
  knitr::kable(caption = "Number and percent of missing values")


count(jur6, is.na(lotsize_sf), category_detail)

```

## Save and write out

```{r save}

jur6 %>%  as.data.frame() %>% 
  count(notinmont, rockflag, gaitflag, othernotmoco)

jur6 %>% 
  count(category)

jur6 <- jur6 %>% 
  select(-c(n.overlaps, origins, validgeo, validgeo.x, validgeo.y,
         ALAND, AWATER, FUNCSTAT, NAMELSAD, GEOID,
         BEGINDATE, BEGINREASO, BEGINPLANN, ENDDATE, ENDREASON,
         ENDPLANNER))

saveRDS(jur6,
        paste0(jdir, 
               "postcleaned-",
               filepath,
               "-data.Rdata"))
```


### Write out clean data.

```{r write}
write_csv(jur6, 
          paste0(jdir, 
               "postcleaned-", filepath, "-data.csv"))
```



