---
title: "Regional Housing Framework"
subtitle: "Pre-clean Fairfax County public records data"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Chris Davis, adapted from Sarah Strochak

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The purpose of this program is to obtain additional data from the county to supplement the Black Knight data when possible. Originally based on preclean_arlington.rmd.

## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)

source("../Macros/read-bk.R")
source("../Macros/filter-bk.R")
source("../Macros/select-vars.R")

```


Set FIPS code, filepath name, and directory for data storage (on L drive)

```{r fips}
currentfips <- "51059"
filepath <- "fairfax"
jdir <- paste0("L:/Libraries/RegHsg/Data/", filepath, "/")
rdir <- paste0("L:/Libraries/RegHsg/Raw/", filepath, "/")

#jdir <- paste0("D:/DCDATA/RegHsg/Data/", filepath, "/")
#rdir <- paste0("D:/DCDATA/RegHsg/Raw/", filepath, "/")


```


Create directory for data exports
```{r dir}
if (!dir.exists("../Data")) {
  dir.create("../Data")
}

if (!dir.exists(paste0("L:/Libraries/RegHsg/Raw/", filepath))) {
  dir.create(paste0("L:/Libraries/RegHsg/Raw/", filepath))
}

# if (!dir.exists(paste0("D:/DCDATA/RegHsg/Raw", filepath))) {
#   dir.create(paste0("D:/DCDATA/RegHsg/Raw", filepath))
# }


```

Load in Black Knight data for the region, select jurisdiction and standard variables
```{r read}
if (!exists("region")) {
  region <- read_bk("dc-cog-assessment_20181228.csv")
} else {
  warning("region data already read in")
}

jur <- region %>% 
  filter_bk(fips = currentfips) %>% 
  select_vars()

```
## Download files

Fairfax has several files of use  
- Tax Administration's Real Estate - Parcel-level file, Buildable/non-buildable, description of property, zone  
- Tax Administration's Real Estate - Land Area file - SF land area, # Units (although seems way too high for some), Type of building  
- Tax Administration's Real Estate - Dwelling Data - Has number of stories of parcel  

```{r download-data}
pafile <- paste0(rdir, filepath, "-parcel-file.csv")
landfile <-paste0(rdir, filepath, "-land-file.csv")
dwellingfile <- paste0(rdir, filepath, "-dwelling-file.csv")

# parcel file
if (!file.exists(pafile)) {
  download.file("https://opendata.arcgis.com/datasets/7607cf5046c5495183251d1c9dba0014_1.csv",
                destfile = pafile)
}
# land file
if (!file.exists(landfile)) {
  download.file("https://opendata.arcgis.com/datasets/f1f0f31844cf49489134f9fa2b8f16f5_3.csv",
                destfile = landfile)
}

# real estate file - dwelling
if (!file.exists(dwellingfile)) {
  download.file("https://opendata.arcgis.com/datasets/53ee1065351c4273ab91ba2e6cfbbc6d_2.csv",
                destfile = dwellingfile)
}

```
## Read files

```{r read-files}

parcel <- read_csv(pafile)

land <- read_csv(landfile)

dwelling <- read_csv(dwellingfile)


rm(pafile, landfile, dwellingfile)

```

## Clean and merge

Select variables from each dataset that we want to keep.

From the parcel file, we will keep the location descriptions and the zoning data in case we need to confirm zoning information and category of the parcels. 

We also archive the raw parcel ID and create a new version that will merge with the Black Knight data.

```{r select-parcel}
parcel_temp<- parcel %>% 
  select(raw_parcelid_par = PARID,
         num_liv_unit = LIVUNIT,
         loc_desc_tax = LOCATION_DESC, 
         luc_desc_tax = LUC_DESC,
         zone_tax = ZONING_DESC) %>% 
  mutate(parcel_id = str_replace_all(raw_parcelid_par, " ", "")) %>%
  filter(!is.na(raw_parcelid_par))

```


There are `r parcel_temp %>% group_by(parcel_id) %>% filter(n()>1) %>% nrow()` parcels that show up more than once on the parcel file. These may be unit counts or portions of parcels, but it is unclear. 

For now, for each parcel that has multiple entries, I take the first value of all of the other values. Can clean up later if we want to use this data
```{r parcel}
parcel_final<-parcel_temp %>% 
  group_by(parcel_id) %>%
  summarize(loc_desc_tax=first(loc_desc_tax),
  luc_desc=first(luc_desc_tax),
  zone_tax = first(zone_tax),
  num_livunit=first(num_liv_unit))
```


(Chris add)
As the parcel IDs are not comprable, we will add a variable with a pin that eliminates the "-"s from the Blacknight file. 

```{r mutate}
jur_temp <- jur %>%
        mutate(parcel_id = str_replace_all(assessorsparcelnumberapnpin, "-", ""))
```

There are `r (!jur_temp$parcel_id%in%parcel_final$parcel_id) %>% sum` units in the Blacknight data that are not in the parcels data.  
There are `r (!parcel_final$parcel_id%in%jur_temp$parcel_id) %>% sum` units in the parcels data that are not in the BlackKnight data.


From the dwelling file, we will keep a variable designating the number of stories of the dwelling.

We also archive the raw parcel ID and create a new version that will merge with the Black Knight data.



```{r select-dwelling}
dwelling_temp <- dwelling %>% 
  select(raw_parcelid_par = PARID,
         stories_tax = STYLE_DESC, 
         possibleSF_tax = SFLA) %>% 
  mutate(parcel_id = str_replace_all(raw_parcelid_par, " ", "")) %>%
  filter(!is.na(raw_parcelid_par))

```

There are `r dwelling_temp %>% group_by(parcel_id) %>% filter(n()>1) %>% nrow()` parcels that show up more than once on the dwelling file.
There are sometimes also multiple stories for the same parcel. If we do use this, I need to go back and clean this. 

There are `r (!jur_temp$parcel_id%in%dwelling_temp$parcel_id) %>% sum` units in the Blacknight data that are not in the dwelling data. 
There are `r (!dwelling_temp$parcel_id%in%jur_temp$parcel_id) %>% sum` units in the dwelling data that are not in the Blacknight data.  



For now, for each parcel that has multiple entries, I take the first value of stories and sum the possible the SF variable. 
```{r summarize-dwelling}
dwelling_final<-dwelling_temp %>% 
  group_by(parcel_id) %>%
  summarize(stories_tax=first(stories_tax), 
            dwellings_per_parcel = n(), 
            possibleSF_tax = sum(possibleSF_tax, na.rm=T))
```




From the land file, we will save land area, number of units, and category of parcel
```{r select-land}
land_temp  <- land %>% 
  select(raw_parcelid_land = PARID,
         area_tax = SF,
         units_tax = UNITS,
         category_tax = CODE_DESC) %>% 
  mutate(parcel_id = str_replace_all(raw_parcelid_land, " ", "")) %>%
  filter(!is.na(raw_parcelid_land))
 
```

There are `r land_temp %>% group_by(parcel_id) %>% filter(n()>1) %>% nrow()` parcels that show up more than once in the land file.  

There are `r (!jur_temp$parcel_id%in%land_temp$parcel_id) %>% sum` units in the Blacknight data that are not in the dwelling data. 
There are `r (!land_temp$parcel_id%in%jur_temp$parcel_id) %>% sum` units in the dwelling data that are not in the Blacknight data.  

For now, for each parcel that has multiple entries, I take the sum of the area and the first value of all of the other values. Can clean up later if we want to use this data.


```{r summarize-land}
land_final<-land_temp %>% 
  group_by(parcel_id) %>%
  summarize(area_tax=sum(area_tax, na.rm=T), 
            land_parcel_sum = n(),
            units_tax = first(units_tax),
            category_tax = first(category_tax))
```



Merge

```{r merge}

jur1 <- left_join(jur_temp, parcel_final, 
                 by = "parcel_id")

jur2 <- left_join(jur1, dwelling_final, 
                 by = "parcel_id")

jur_final <- left_join(jur2, land_final, 
                 by = "parcel_id")

```

## Save

Since this is an intermediary dataset, save as an R dataset for easy reading into the the next step in the cleaning.

```{r save}

saveRDS(jur_final,
        paste0(jdir, 
               "precleaned-",
               
               filepath,
               "-data.Rdata"))

```
