---
title: "Regional Housing Framework"
subtitle: "Vacant lots analysis for Arlington County: special exceptions"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Sarah Strochak

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The goal of this analysis is to modify the previous vacant lots analysis to include special exceptions that are written into the zoning code. We will rely on two datasets that were created in `vacant-lots-arlington.Rmd` - the cleaned general zoning code, and the cleaned set of vacant lots.

The site plan process is mostly used for multifamily construction, so we will consider only the special exceptions for multiple family dwellings.


## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)
library(urbnthemes)

set_urbn_defaults("print")

```

Create directory for data exports
```{r dir}

if (!dir.exists("L:/Libraries/RegHsg/Prog/Tables")) {
  dir.create("L:/Libraries/RegHsg/Prog/Tables")
}
```

Set FIPS code and filepath name
```{r fips}

currentfips <- "51013"
filepath <- "arlington"


```

Load in cleaned vacant lots and zoning code.

```{r read}

vacant <- readRDS("L:/Libraries/RegHsg/Data/arlington/arlington-vacant-lots.Rdata")
zoningcode <- readRDS("L:/Libraries/RegHsg/Data/arlington/arlington-clean-zoning-code.Rdata") %>% 
  # create flag to indicate these do not come from the special exceptions file
  mutate(special_exception = 0)

```


## Clean special exceptions.

### Move file 

First- **manually** move file from Box to the `zoning-codes` directory in the `Doc` folder on the L drive.

### Read file

```{r read-zoning-csv}

z <- c("text", "text", "text", "numeric",
       "numeric", "numeric", "text", "text",
       "numeric", "numeric", "numeric", "text",
       "text", "numeric", "numeric", "numeric",
       "numeric", "text", "numeric")


se <- readxl::read_excel("L:/Libraries/RegHsg/Doc/zoning-codes/Arlington Multiple Family Special Exceptions 2.19.19.xlsx", 
                         range = "A1:S49", 
                         col_types = z)

names(se) <- 
  c("zoning_code", "district_type", "dwelling_type",
    "site_area", "lot_area_min", "lotarea_per_unit",
    "lot_width", "height_ft", "height_stories",
    "lot_coverage_max", "floor_area_min", "far_max",
    "setbacks_center", "setbacks_row",
    "front_side_yard", "side_yard", "frontage", "comments", "density_units_per_acre")

rm(z)

```

### Clean up special exceptions

A few cleaning tasks:

* We need to extract the oddities in the height variable, and archive them in case we need them later. We reset these to NA, and convert the variable type back to numeric.
* Create `special_exception` flag
* Filter out `dwelling_type` values that we cannot use
* Change `dwelling_type` values incorrectly recorded as "Multiple-family dwellings" to "Multiple-family"

```{r height-problems}

height_problems <- se %>% 
  filter(str_detect(height_ft, "[:alpha:]"))

se1 <- se %>% 
  mutate(height_ft = ifelse(str_detect(height_ft, "[:alpha:]") == TRUE,
                            NA,
                            as.numeric(height_ft))) %>% 
  mutate(special_exception = 1) %>% 
  mutate(dwelling_type = ifelse(dwelling_type == "Multiple-family dwellings",
                                "Multiple-family",
                                dwelling_type)) %>% 
  filter(dwelling_type != "Hotel")
  
  
```

Check to make sure there are no codes in the special exceptions that are NOT in the by-right zoning code, due to being entered incorrectly.

```{r check-se-code}

se1 %>% 
  filter(!zoning_code %in% zoningcode$zoning_code) %>% 
  nrow()

```
Filter out multiple-family exceptions.

```{r get-mf}

semf <- se1 %>% 
  filter(dwelling_type == "Multiple-family")

```

### Fill in missing variables in special exceptions

If the special exceptions inherit the same stipulations as the by-right zoning code, the fields were left blank. Use `fill()` to fill in these gaps.

Replace missing lot coverage maximums with 56%- this is to be used if there is no specific designation, per the Arlington County zoning code.

```{r fil}

mf_fill <- semf %>% 
  bind_rows(filter(zoningcode,
                   zoning_code %in% semf$zoning_code,
                   dwelling_type == "Multiple-family")) %>% 
  group_by(zoning_code, dwelling_type) %>% 
  fill(lot_area_min, .direction = "down") %>% 
  fill(lot_area_min, .direction = "up") %>% 
  fill(lotarea_per_unit, .direction = "down") %>% 
  fill(lotarea_per_unit, .direction = "up") %>%
  fill(height_ft, .direction = "down") %>% 
  fill(height_ft, .direction = "up") %>% 
  fill(lot_coverage_max, .direction = "down") %>% 
  fill(lot_coverage_max, .direction = "up") %>% 
  mutate(lot_coverage_max = ifelse(is.na(lot_coverage_max),
                                   .56,
                                   lot_coverage_max))

```

There are four zoning codes that only allow for multiple-family development via special exception. The stipulations vary based on the 

56 percent if no coverage ratio
create two height variables and take the lower
FAR / 

# how many observations are the same?
```{r }

se_compare %>% nrow() -
se_compare %>% 
  distinct(lot_area_min, lotarea_per_unit, height_ft, lot_coverage_max) %>% 
  nrow()

se_compare %>% 
  group_by(dwelling_type) %>% 
  count()

```

Except for multiple-family, the special exceptions, within a dwelling_type, were the same as the by-right zoning codes. For many zoning codes, additional dwelling types were added. 

### Dwelling type hierarchy

We can adjust the hierarchy we used to select the most dense zoning codes in the previous analysis by adding those that now occur, and prioritizing the multiple-family dwelling type from the special exception over the by-right. 

First, remove dwelling types that are not possible to include.

```{r hier-dwell}

zoningcode_se <- se_compare %>% 
  ungroup() %>% 
  filter(dwelling_type != "Low or Moderate Income Housing") %>% 
  filter(!(dwelling_type %in% c("All",
                                "All Other",
                                "Duplex; share a lot line with RA, C, or M",
                                "Semi-detatched; share a lot line with RA, C, or M"))
           & n() > 1) %>% 
  mutate(dwelling_type = ifelse(dwelling_type %in% c("Multiple-family", "Apartment"),
                                "Multiple-family special exception",
                                dwelling_type)) %>% 
  mutate(dwelling_type = factor(dwelling_type,
                                levels = c("Multiple-family special exception",
                                           "Multiple-family",
                                           "Townhouse",
                                           "Semi-detached",
                                           "Duplex",
                                           "Other Residential",
                                           "One Family Dwelling")),
         one_unit_flag = ifelse(dwelling_type == "One Family Dwelling",
                                1,
                                0),
         two_unit_flag = ifelse(dwelling_type == "Duplex",
                                1,
                                0))

```

