---
title: "Regional Housing Framework"
subtitle: "Vacant lots analysis for Arlington County"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Sarah Strochak

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The goal of this analysis is to determine how many units could be added by filling vacant lots with the most-dense developments allowed under the current zoning codes.
The two steps are:

1. Merge cleaned vacant lots data with zoning code data.
2. Calculate how many units are allowed one each vacant lot.

## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)
library(urbnthemes)

set_urbn_defaults("print")

source("../Macros/read-jurisdiction.R")

```

Create directory for data exports
```{r dir}
if (!dir.exists("../Data")) {
  dir.create("../Data")
}

```

Set FIPS code and filepath name
```{r fips}
currentfips <- "51013"
filepath <- "arlington"
```

Load in Black Knight data for the region, select jurisdiction and standard variables
```{r read}
if (!exists("jur")) {
  jur <- read_jurisdiction(filepath)
} else {
  warning(filepath, " data already read in")
}

```

Select only vacant lots

```{r res}

vacant <- jur %>% 
  filter(category == "vacant")
  
```

## Detailed look at vacant lots

See how many lots have missing key variables

```{r missing}

vacant %>% 
  group_by(category_detail) %>% 
  summarize_at(vars(propaddress, lotsize_sf, zoning),
               ~ sum(is.na(.))) %>% 
  mutate_at(vars(propaddress, lotsize_sf, zoning),
            .funs = funs(percent = scales::percent(. / nrow(vacant)))) %>% 
  knitr::kable(caption = "Number and percent of missing values")

```

### Missing zoning variables


## Clean zoning code information

### Move file 

First- **manually** move file from Box to the `zoning-codes` directory in the `Doc` folder on the L drive. Name file accordingly:
"jursidiction-zoning-code.csv"

### Read file

Read in zoning code, rename variables so that they are machine readable

```{r file-check}

if (!file.exists(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              "-zoning-code.csv"))) {
  
  stop("Move zoning file for jurisdiction to L:/Libraries/RegHsg/Doc/zoning-codes directory")
}

```

```{r read-zoning}
z <- cols(
  Code = col_character(),
  `District Type` = col_character(),
  `Type of Dwelling` = col_character(),
  `Site Area` = col_number(),
  `Lot area` = col_number(),
  `Lot area per dwelling unit (sq. ft.) (minimum)` = col_number(),
  `Lot Width (average minimum, ft)` = col_character(),
  `Height (ft)` = col_double(),
  `Height (stories)` = col_double(),
  `Lot coverage maximum` = col_number(),
  `Floor area minimum` = col_double(),
  `Maximum Density of Floor-Area-Ratio (FAR)` = col_character(),
  `Setbacks (centerline)` = col_character(),
  `Setbacks (right of way line)` = col_double(),
  `Front/Side yard (ft)` = col_double(),
  `Side yard (both sides: minimum percent of required width)` = col_number(),
  Frontage = col_double(),
  Comments = col_character()
)

zoningcode <- read_csv(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              "-zoning-code.csv"),
                       col_types = z)

names(zoningcode) <- 
  c("zoning_code", "district_type", "dwelling_type",
    "site_area", "lot_area_min", "lotarea_per_unit",
    "lot_width", "height_ft", "height_stories",
    "lot_coverage_max", "floor_area_min", "far_max",
    "setbacks_center", "setbacks_row",
    "front_side_yard", "side_yard", "frontage", "comments")

# fix variables read in as whole numbers (should be percent)
zoningcode <- zoningcode %>% 
  mutate_at(vars(lot_coverage_max, side_yard),
            ~ . / 100)
rm(z)

```

### Creating dwelling_type hierarchy

For the vacant lots analysis, we want to determine the `dwelling_type` that allows for the most dense construction for that zone. 

First, we look at all the options, and filter out all the observations marked as "other", since we cannot be sure what property type that refers to. We will also filter out categories that are dependent on adjacent lot lines.

```{r tab-dwellings}

zoningcode %>% 
  count(dwelling_type) %>% 
  knitr::kable()

zoningcode <- zoningcode %>% 
  filter(!dwelling_type %in% c("All",
                               "All Other",
                               "Duplex; share a lot line with RA, C, or M",
                               "Semi-detatched; share a lot line with RA, C, or M"))

```
Since the variables are not always comparable, the following order, from most to least dense, applies:

1. Multiple-family
2. Duplex
3. Townhouse
4. Semi-detached
5. Other Residential
6. One Family Dwelling

### Create factor variable
To use this hierarchy, we transform the `dwelling_type` variable into a factor, assign the factors to the appropriate level, (1-6, as listed above).

```{r dwelling-factor}

zoningcode <- zoningcode %>% 
  mutate(dwelling_type = factor(dwelling_type,
                                levels = c("Multiple-family",
                                           "Duplex",
                                           "Townhouse",
                                           "Semi-detached",
                                           "Other Residential",
                                           "One Family Dwelling")))

# vacant_zoning <- zoningcode %>% 
#   group_by(zoning_code) %>% 
#   slice(which.min(dwelling_type)) 

```

## Merge with cleaned data

Now that we have selected the correct records from the zoning code file, we can merge with the cleaned property records data. First, we have to do a bit of cleaning of these codes, including getting rid of instances where there is a tie, and correcting some of the notation. We will create a new variable in order to archive the original `zoning` variable.

```{r bk-clean}
vacant <- vacant %>% 
  mutate(zoning_code = ifelse(str_detect(zoning, "/") == TRUE,
                              NA,
                              zoning),
         zoning_code = ifelse(zoning_code == "character(0)",
                              NA,
                              zoning_code),
         zoning_code = ifelse(zoning_code == "RC",
                              "R-C",
                              zoning_code),
         zoning_code = ifelse(zoning_code == "C-O-ROSSLYN",
                              "C-O Rosslyn",
                              zoning_code),
         zoning_code = ifelse(zoning_code == "RA-H-3.2",
                              "RA-H 3.2",
                              zoning_code))
```


Merge on zoning code variable. Test that all variables merged.

```{r zoning-merge}

zoningdata <- left_join(vacant, zoningcode, by = "zoning_code")

```

There are `r ` observations that did not have a match in the zoning code. Below, we see which zoning codes were int he Black Knight data but NOT in the zoning code data.

```{r antijoin}
#
#nrow(anti_join(vacant, vacant_zoning, by = "zoning_code"))
#anti_join(vacant, vacant_zoning, by = "zoning_code") %>% 
#  count(zoning_code) %>% 
#  knitr::kable()

```

Aside from those with missing zoning information, these zoning codes do not allow for residential properties, so they could not be found in the zoning code data.

Next, we can take a closer look at properties that have missing zoning information and see if we can figure out why.

```{r miss-zoning}
missing_zoning <- vacant %>% 
  filter(is.na(zoning_code))

missing_zoning %>% 
  summarize(missing_address = sum(is.na(propaddress)),
            zoning_tie = sum(str_detect(zoning, "/") == TRUE, na.rm = TRUE))

```

The missing zoning designations can be explained by either a missing address or multiple zoning codes indicated in the Black Knight data.

```{r other-missing}
missing_zoning %>% 
  filter(!is.na(propaddress),
         str_detect(zoning, "/") == FALSE) %>% 
  nrow()

```

Filter out lots that are too small to build on, select most dense zoning code for those that are big enough
```{r filter-small}

zoningdata1 <- zoningdata %>% 
  filter(!is.na(lot_area_min),
         lotsize_sf >= lot_area_min) %>% 
  group_by(propaddress) %>% 
  slice(which.min(dwelling_type)) %>% 
  ungroup()

```


## Calculate missing units

There are several variables we will use to calculate how many units we can put on each vacant lot.

1. Buildable area: this is `lotsize_sf` x the coverage ratio (`lot_coverage_max`)
2. Living area: this is the buildable area x the construction efficiency. The lower bound estimate is 70%, the upper bound estimate is 85%.
3. Units per floor: this is the living area divided by the average unit size. We use 900 sq. feet, based on the American Housing Survey 2017 data for the DC metro area.
4. Units allowed based on height limits: This is the units per floor x the number of stories.
3. Units allowed based on lot size: this is the lot area divided by `lotarea_per_unit`

We can create a function that will allow us to easily adjust these parameters and compare results.

```{r get-est}

get_unit_variables <- function(dataset, 
                               single_designation, 
                               construction_efficiency, 
                               average_unit_size) {
  
  dataset %>% 
    mutate(parameters = paste0("CE: ", construction_efficiency * 100,
                               "% , Unit size: ", average_unit_size,
                               " sf"),
           buildable_area = lotsize_sf * lot_coverage_max,
           living_area = buildable_area * construction_efficiency,
           units_per_floor = floor(living_area / average_unit_size),
           floors = floor(height_ft / 10),
           units_height = units_per_floor * floors,
           units_lotsize = floor(lotsize_sf / lotarea_per_unit)) %>% 
    mutate_at(vars(units_height, units_lotsize),
              funs(ifelse(dwelling_type == single_designation, 
                          1, 
                          .))) %>% 
    mutate(units_total = case_when(
      dwelling_type == "One Family Dwelling" ~ 1,
      units_height < units_lotsize ~ units_height,
      TRUE ~ units_lotsize))
  
}

get_unit_summary <- function(dataset, 
                             single_designation, 
                             construction_efficiency, 
                             average_unit_size,
                             dwelling_type_breakdown) {

  x <- dataset %>% 
    get_unit_variables(single_designation, 
                       construction_efficiency,
                       average_unit_size) %>%
    group_by(parameters) %>% 
    summarize(lots = n(),
              units_height = sum(units_height),
              units_lotsize = sum(units_lotsize),
              units_total = sum(units_total),
              units_per_lot = floor(sum(units_total / lots)))
  
  y <- dataset %>% 
    get_unit_variables(single_designation, 
                       construction_efficiency,
                       average_unit_size) %>%
    group_by(parameters, dwelling_type) %>% 
    summarize(lots = n(),
              units_height = sum(units_height),
              units_lotsize = sum(units_lotsize),
              units_total = sum(units_total),
              units_per_lot = floor(sum(units_total / lots))) %>% 
    mutate(dwelling_type = as.character(dwelling_type))
  
  
  if (dwelling_type_breakdown == TRUE) {
   
      bind_rows(y, mutate(x, dwelling_type = "All"))  
  
  } else if (dwelling_type_breakdown == FALSE) {
      x
    }
}



```
    

Get summary statistics for several different parameters

```{r sum}

get_unit_summary(zoningdata1,
                 "One Family Dwelling",
                 .7,
                 700,
                 dwelling_type_breakdown = TRUE) %>% 
  knitr::kable()


map_df(c(.7, .85), ~ get_unit_summary(zoningdata1,
                                      "One Family Dwelling",
                                      .,
                                      700,
                                      dwelling_type_breakdown = FALSE)) %>% 
  knitr::kable()

```

Look at the distribution of units for lots with more than 1 unit, to check for extreme outliers.

```{r scatter, echo = FALSE}

zoningdata1 %>% 
  get_unit_variables("One Family Dwelling", .7, 700) %>% 
  filter(units_total > 1) %>% 
  ggplot() +
  geom_point(mapping = aes(lotsize_sf, units_total),
             alpha = .5) +
  scale_x_continuous(labels = scales::comma,
                     expand = expand_scale(mult = c(0, 0.002)), 
                     limits = c(0, 300000),
                     breaks = 0:6 * 50000) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.002)), 
                     limits = c(0, 300),
                     breaks = 0:7 * 50) +
  scatter_grid() +
  labs(x = "Lot size (square feet)", y = "Total units") +
  theme(plot.margin = margin(t = 20, r = 20))

```


Look more closely at vacant lots < 50000 SF

```{r scatter-small, echo = FALSE}

zoningdata1 %>% 
  get_unit_variables("One Family Dwelling", .7, 700) %>% 
  filter(lotsize_sf < 50000,
         units_total > 1) %>% 
  ggplot() +
  geom_point(mapping = aes(lotsize_sf, units_total,
                           color = dwelling_type),
             alpha = .8) +
  scale_x_continuous(labels = scales::comma,
                     expand = expand_scale(mult = c(0, 0.0002)), 
                     limits = c(0, 50000),
                     breaks = 0:5 * 10000) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.0002)), 
                     limits = c(0, 60),
                     breaks = 0:4 * 15) +
  scatter_grid() +
  labs(x = "Lot size (square feet)", y = "Total units") +
  theme(plot.margin = margin(t = 20,
                             r = 20))



```