---
title: "Regional Housing Framework"
subtitle: "Vacant lots analysis for DC"
author: "YS"
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Yipeng adapted from Sarah Strochak

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The goal of this analysis is to determine how many units could be added by filling vacant lots with the most-dense developments allowed under the current zoning codes.
The two steps are:

1. Merge cleaned vacant lots data with zoning code data.
2. Calculate how many units are allowed one each vacant lot.

## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)
library(urbnthemes)
library(sf)

set_urbn_defaults("print")

```

Create directory for data exports
```{r dir}
if (!dir.exists("../Data")) {
  dir.create("../Data")
}

if (!dir.exists("L:/Libraries/RegHsg/Prog/Tables")) {
  dir.create("L:/Libraries/RegHsg/Prog/Tables")
}
```

Set FIPS code and filepath name
```{r fips}

currentfips <- "11001"
filepath <- "DC"
```


Load in cleaned Black Knight data for `r str_to_title(filepath)` County.

```{r read}

read_dc <- function(rmd = TRUE) {
  
  filename <- paste0("L:/Libraries/RegHsg/Data/", filepath,
                     "/", filepath, "-cleaned-data.csv")
  
  if (!file.exists(filename)) {
    stop("cleaned data not found in Data directory")
  } else {
    
    read_csv(filename,
             col_types = cols(county_fips = col_character(),
                              county_name = col_character(),
                              parcelbase_SSL = col_character(),
                              propaddress = col_character(),
                              propcity = col_character(),
                              propstate = col_character(),
                              propzip = col_character(),
                              propunitno = col_character(),
                              prophouseno = col_character(),
                              propstreetname = col_character(),
                              propstreetsuffix = col_character(),
                              lat = col_double(),
                              long = col_double(),
                              tract = col_character(),
                              owneroccupiedresidential = col_character(),
                              countylandusedescription = col_character(),
                              zoning = col_character(),
                              buildingarea = col_double(),
                              noofbuildings = col_character(),
                              noofstories = col_character(),
                              numberofunits = col_integer(),
                              yearbuilt = col_integer(),
                              lotsize_acres = col_double(),
                              lotsize_sf = col_double(),
                              address_type = col_character(),
                              category = col_character(),
                              category_detail = col_character(),
                              residential = col_integer(),
                              building_type = col_character(),
                              parcelbase_address = col_character(),
                              parcelbase_usecode = col_character(),
                              parcelgeo_x = col_double(),
                              parcelgeo_y = col_double(),
                              vacant_flag = col_integer()))
  }
}


if (!exists("jur")) {
  jur <- read_dc()
} else {
  warning(filepath, " data already read in")
}


```

## Clean zoning code information

### Move file 

First- **manually** move file from Box to the `zoning-codes` directory in the `Doc` folder on the L drive. Name file accordingly:
"jursidiction-zoning-code.csv"

### Read file

Read in zoning code, rename variables so that they are machine readable

```{r file-check}

if (!file.exists(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              " zoning code.csv"))) {
  
  stop("Move zoning file for jurisdiction to L:/Libraries/RegHsg/Doc/zoning-codes directory")
}

```


```{r read-zoning-csv}
z <- cols(
   `Code` = col_character(),
  `District Type` = col_character(),
  `Type of Dwelling` = col_character(),
  `Number of Dwelling units allowed` = col_number(),
  `Site Area` = col_number(),
  `Lot area` = col_number(),
  `Lot area per dwelling unit (sq. ft.) (minimum)` = col_number(),
  `Lot Width (average minimum, ft)` = col_character(),
  `Height (ft)` = col_double(),
  `Height (stories)` = col_double(),
  `Lot coverage maximum` = col_number(),
  `Floor area ratio max`=col_double(),
  `Maximum Density of Floor-Area-Ratio (FAR)` = col_character(),
  `Setbacks (centerline)` = col_character(),
  `Setbacks (right of way line)` = col_double(),
  `Side yard (ft)` = col_number(),
  `Side yard (both sides: minimum percent of required width)` = col_number(),
  `Rear yard` = col_double(),
  `Frontage` = col_double(),
  `Comments` = col_character()
)


zoningcode <- read_csv(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              " zoning code.csv"),
                       col_types = z)

names(zoningcode) <- 
  c("zoning_code", "district_type", "dwelling_type","dwelling_units_allowed",
    "site_area", "lot_area_min", "lotarea_per_unit",
    "lot_width", "height_ft", "height_stories",
    "lot_coverage_max","floor_area_ratio_max", "far_max",
    "setbacks_center", "setbacks_row",
    "front_side_yard", "side_yard_bothside","rear_side_yard", "frontage", "comments")

# fix variables read in as whole numbers (should be percent)
zoningcode <- zoningcode %>% 
  mutate_at(vars(lot_coverage_max, side_yard_bothside),
            ~ . / 100)
rm(z)

```

### Creating dwelling_type hierarchy

For the vacant lots analysis, we want to determine the `dwelling_type` that allows for the most dense construction for that zone. 

First, we look at all the options, these categories are not as helpful because attached and detached can have different meaning in the context ofdifferent zoning types (both residential house and residential flat has these two types, for example)plus we have a lot of NAs for residential code. So need to create a new dwelling type variable.

```{r tab-dwellings}
zoningcode %>% 
  count(dwelling_type) %>% 
  knitr::kable()

```

Count the number of unique zoning codes to make sure we won't get rid of any codes entirely.

```{r count-zoning}
length(unique(zoningcode$zoning_code))

```


Plus, the DC dwelling type is missing for a lot of zoning, so it is difficult to follow the exact same method. I will here create a new dwelling type variable based on #of units allowed, district type variable. 

Since the variables are not always comparable, after the recoding, from most to least dense, order applies:

1. RA
2. RF4 
3. RF3
4. RF2 
5. One Family 

Then within each category, DC has different requirements for each different type of dwelling within the same zoning code. So need to furthur break down the new dwelling type variable. 

```{r recode dwelling}
zoningcode1 <- zoningcode %>% 
  mutate(newdwelling_type = case_when(
    dwelling_units_allowed==1 & lot_area_min>=5000~ "One Family-A",
    dwelling_units_allowed==1 & dwelling_type== "detached"~ "One Family-A",
    dwelling_units_allowed==1 & dwelling_type== "IZ detached"~ "One Family-B",
    dwelling_units_allowed==1 & dwelling_type== "semi-detached"~ "One Family-C",
    dwelling_units_allowed==1 & dwelling_type== "attached"~ "One Family-D",
    dwelling_units_allowed==1 & dwelling_type== "IZ attached"~ "One Family-E",
    dwelling_units_allowed==1 & dwelling_type== "IZ semi-detached"~ "One Family-F",
    dwelling_units_allowed==1 & dwelling_type== "all other structures"~ "All Other",
    dwelling_units_allowed==2 & dwelling_type== "semi-detached" ~ "RF2-A",
    dwelling_units_allowed==2 & dwelling_type== "row dwelling or flat" ~ "RF2-B",
    dwelling_units_allowed==2 & dwelling_type== "row dwelling or flat (IZ)" ~ "RF2-C",
    dwelling_units_allowed==2 & dwelling_type== "all other structures" ~ "All Other",
    dwelling_units_allowed==3 & dwelling_type=="semi-detached" ~ "RF3-A",
    dwelling_units_allowed==3 & dwelling_type=="row dwelling or flat" ~ "RF3-B",
    dwelling_units_allowed==3 & dwelling_type=="row dwelling or flat (IZ)" ~ "RF3-C",
    dwelling_units_allowed==3 & dwelling_type=="all other structures" ~ "All Other",
    dwelling_units_allowed==4 & dwelling_type=="semi-detached"~ "RF4-A",
    dwelling_units_allowed==4 & dwelling_type=="row dwelling or flat"~ "RF4-B",
    dwelling_units_allowed==4 & dwelling_type=="row dwelling or flat (IZ)"~ "RF4-C",
    dwelling_units_allowed==4 & dwelling_type=="all other structures"~ "All Other", 
    floor_area_ratio_max>1 & is.na(dwelling_units_allowed) ~ "RA",
    is.na(floor_area_ratio_max) &is.na(dwelling_units_allowed) ~ "All Other",
    floor_area_ratio_max <= 1 ~ "Other Mixed",
    zoning_code=="RA-1"|zoning_code=="RA-6" ~ "RF2-B"  
  ))
#RA-1 and RA-6 are most comparable to RF2-B according to zoning documentation

```
Filter out dwelling_types of "All", "All Other", and those that are dependent on a shared boundary.

```{r purge}
zoningcode2 <- zoningcode1 %>% 
  group_by(zoning_code) %>% 
  filter(!(newdwelling_type %in% c("All Other",
                               "Other Mixed")
           & n() > 1)) %>% 
  ungroup()

# I could put IZ-detached IZ semi detached etc into other mixed and exclude from the merge?

```

 Make sure we did not lose any zoning codes entirely.
```{r responsible-purge}

stopifnot(length(unique(zoningcode$zoning_code))
          == length(unique(zoningcode1$zoning_code)))

```

### Create factor variable
To use this hierarchy, we transform the `newdwelling_type` variable into a factor, assign the factors to the appropriate level, (1-6, as listed above).
We will also create a designation for dwelling types that are restricted to a certain amount of units (one unit for "One Family Dwelling" and one for two units for "Duplex"). 

**NOTE** This will be VERY jurisdiction specific.

**NOTE** If there are no one or two unit limitations, still make the variable and set them all equal to zero. This allows the functionality of the calculation to be more uniform across counties.

```{r dwelling-factor}

zoningcode_factor <- zoningcode2 %>% 
  mutate( newdwelling_type = factor(newdwelling_type,
                                levels = c("RA",
                                           "RF4-C","RF4-B", "RF4-A",
                                           "RF3-C","RF3-B","RF3-A",
                                           "RF2-C","RF2-B","RF2-A",
                                           "One Family-F", "One Family-E","One Family-D", "One Family-C", "One Family-B", "One Family-A")),
         one_unit_flag = ifelse(dwelling_units_allowed==1, 1,0),
         two_unit_flag = ifelse(dwelling_units_allowed==2, 1,0),
         three_unit_flag = ifelse(dwelling_units_allowed==3, 1,0),
         four_unit_flag = ifelse(dwelling_units_allowed==4, 1,0))


```

## Clean vacant lots data

Select only vacant lots

```{r res}

vacant <- jur %>% 
  filter(category == "vacant")
  
```

See how many lots have missing key variables

```{r missing}
vacant %>% 
  mutate(num = NA) %>% 
  group_by(category_detail) %>% 
  summarize_at(vars(propaddress, lotsize_sf, zoning, num),
               ~ sum(is.na(.))) %>% 
  mutate_at(vars(propaddress, lotsize_sf, zoning),
            .funs = list(percent = ~scales::percent(. / num))) %>% 
  knitr::kable(caption = "Number and percent of missing values")

```

### Missing lot size
We filled in the missing lot size, when we can, with the cleaned parcelbase lot area data in the clean DC step. So skip this step.

We have 788 lotsize missing after the cleaning. 

```{r count-lots}
sum(is.na(vacant$lotsize_sf))
```

### Assessment values

The difference between the land value, improvement value, and total value can help us determine if lots are really vacant.

```{r check-assess}
#this part wont knit not sure why, I'm pasting the table in the comment
#vacant %>% 
  #group_by(category_detail) %>% 
  #summarize(nprops = n(),
            #imp_zero = sum(assessedimprovementvalue == 0, na.rm = TRUE),
           # imp_na = sum(is.na(assessedimprovementvalue))) %>% 
 # mutate(imp_none = imp_zero + imp_na,
       #  percent_none = imp_none / nprops) %>% 
  #knitr::kable()



#|category_detail        | nprops| imp_zero| imp_na| imp_none| percent_none|
#|:----------------------|------:|--------:|------:|--------:|------------:|
#|Vacant unimproved      |  11565|    11385|    180|    11565|            1|
#|Vacant with structures |    242|      242|      0|      242|            1|
  
```

All of the vacant lots in DC have either a 0 or a missing value for the improvement value, which is what we would expect. 

```{r not-so-vacant}

vacantstructure <- vacant%>% 
  filter(category_detail=="Vacant with structures") %>% 
  group_by(countylandusedescription) %>% 
  summarize(n=n()) 

```

We will not include vacant with structures (242 cases) in the calculation. We can also make a flag for missing coordinates.

```{r not-vacant-recat}

vacant1 <- vacant %>% 
  filter(category_detail == "Vacant unimproved") %>% 
  mutate(missing_coord = ifelse(is.na(lat) | is.na(long),
                                1,
                                0),
         parcel_x= parcelgeo_x,
         parcel_y= parcelgeo_y,
         missing_parcelXY= ifelse(is.na(parcelgeo_x) | is.na(parcelgeo_y),
                                1,
                                0)) 
```

### Missing zoning variables
Blackknight zoning data for DC is outdated, making it complicated to crosswalk it to the new zoning code and our zoning statistics. Instead, we will use a spatial join to the county's zoning layer to get zoning information for all parcels. 

Read in zoning layer, set CRS
```{r read-zoning}
zdir <- paste0("L:/Libraries/RegHsg/Maps/",
               filepath,
               "-zoning-map")
zfile <- str_sub(list.files(zdir), end = -5) %>% unique()

zon <- read_sf(dsn = zdir,
               layer = zfile)

st_geometry(zon)
zon <- st_transform(zon, crs = 4326)

```

Convert vacant lots to SF, set CRS. To convert to SF, we need to use only the observations with non-missing longitude and latitude. We separate these off and recombine after the spatial join.

```{r convert-sf}
vacantsf <- vacant1 %>% 
  filter(missing_coord == 0) %>% 
  st_as_sf(coords = c("long", "lat")) %>% 
  st_set_crs(st_crs(zon))

#convert the parcelbase XY to lat and lon
vacantsfparcel <- vacant1 %>% 
  filter(missing_parcelXY == 0) %>% 
  st_as_sf(coords = c("parcel_x", "parcel_y")) %>% 
  st_set_crs(2804) %>% 
  st_transform(4326)

#export for ArcGIS check
vacant1 %>% 
  filter(missing_parcelXY == 0) %>% 
  write_csv(paste0("L:/Libraries/RegHsg/Prog/Tables/",
                   filepath,
                   "DCparcelbase_XY.csv"))

ggplot()+
  geom_sf(data = vacantsfparcel, mapping = aes(),
          color = "black", size=0.1)

```

Plot to make sure they are overlapping
```{r make-map, echo=FALSE}
ggplot() +
  geom_sf(data = zon, mapping = aes(fill = ZONE_DISTR)) +
  scale_fill_manual(values = c("#1696d2", "#fdbf11", 
                               "#d2d2d2", "#ec008b", 
                               "#55b748", "red", "green", "yellow", "blue")) +
  geom_sf(data = vacantsfparcel, mapping = aes(),
          color = "black", size=0.1)

```

Spatial join

```{r spatial-join}

vacantjoin <- st_join(vacantsfparcel, zon, join = st_intersects)

```

After the join, we convert back to a dataframe, for ease of computations.
```{r back-to-df}

vacant2 <- st_set_geometry(vacantjoin, NULL)

```

Count how many observations did not join.(0) This will throw an error if more than 5% did not match- in this case, check the underlying shapefiles for compatability.

```{r count-spatial}

paste0(sum(is.na(vacant2$ZONING_LAB)), " observations didn't join- ",
      round(sum(is.na(vacant2$ZONING_LAB)) / nrow(vacant2) * 100, 2), "%")

if (sum(is.na(vacant2$ZONING_LAB)) / nrow(vacant2) > .05) {
  stop("More than 5% of observations could not be spatially joined.")
}
      
```

Test how many have different results from the Black Knight zoning variable. 3804 of 8661 were the same if using new zoning code. 5448 of 8661 were the same if using old zoning code in the shapefile.This is expected because blacknight data zoning code was dated and not consisted with the 2016 updated DC zoning codes.

```{r test-spatial}

vacant2 %>% 
  group_by(is.na(zoning)) %>% 
  count(zoning == ZONING_LAB)

vacant2 %>% 
  group_by(is.na(zoning)) %>% 
  count(zoning == OLD_ZONING)

```

I'm skipping the comparing blacknight zoning with shapefile zoning because 1. We can only compare the old zoning data. 2. The crosswalk is one directional-- there's no way we can infer the new zoning code from old one, other than the spatial join. So I'm going to use zoning code from the spatial join and merge it with the zoning statistics spreadsheet. 

Recombine with observations with missing coordinates, check that number of rows is the same.

```{r get-final}

vacanttotal <- vacant1 %>% 
  filter(is.na(parcel_x) | is.na(parcel_y)) %>% 
  mutate(zoning_bk = zoning) %>% 
  bind_rows(vacant2)

stopifnot(nrow(vacanttotal) == nrow(vacant1))

```

### Dual zoning codes

The last step is to identify a zoning designation from shapefile that has more than 1 code. There's no dual zoning code from the shapefile. 

```{r real-final}

dual <- vacanttotal %>% 
  filter((str_detect(ZONING_LAB, ",") == TRUE)) 

```

## Merge with cleaned data

```{r zoning-merge}

vacantfinal <- vacanttotal %>% 
        mutate(zoning_code=ZONING_LAB)

zoningdata <- left_join(vacantfinal, zoningcode_factor, by = "zoning_code")

zoningdata %>% 
  group_by(newdwelling_type) %>% 
  count() %>% 
  knitr::kable()

```

There are `r ` observations that did not have a match in the zoning code. Below, we see which zoning codes were in the shapefile data but NOT in the zoning code data.We need to take out the dashes in the zoning code. 

```{r antijoin}

anti_join(vacantfinal,zoningcode_factor, by = c("zoning_code")) %>% 
  count(ZONING_LAB) %>% 
  knitr::kable()

```

We need to clean the spatial joined zoning code, especially the Residential codes started ones that have a dash in the middle that are not joined with the zoning spreadsheet. After cleaning, use anti_join again to test if there are any left over residential codes that are not joined.

```{r remove dash}

vacantcleaned <- vacantfinal %>% 
          mutate(zoning_code.shape=zoning_code) %>% 
          mutate(firsttwo= substr(zoning_code.shape, 1,2),
                 badrescode=ifelse(str_detect( firsttwo, "R-")==1, 1, 0),
                 zoning_code=ifelse(badrescode==1, paste0(substr(zoning_code.shape, 1,1),substr(zoning_code.shape,3,3)),zoning_code.shape))

zoningdatanew <- left_join(vacantcleaned, zoningcode1, by = "zoning_code")

```

After cleaning, use anti_join again to test if there are any left over residential codes that are not joined. R1 and RC-1, RC-3 not joined. Check the spreasheet and they are not in. Leave for now.
```{r antijoinagain}

anti_join(vacantcleaned, zoningcode_factor, by = c("zoning_code")) %>% 
  count(zoning_code) %>% 
  knitr::kable()

```

There are 2026 R1s that are not joined which is a lot, in our spreadsheet we have R1A and R1B, pull in lotsize and see which ones we should assign the R1 to. R1A has lot size of 7500, R1B has lot size of 5000. Use as threshold.
```{r assign R1}

anti_join(vacantcleaned, zoningcode_factor, by = c("zoning_code")) %>% 
  filter(zoning_code=="R1") %>% 
  mutate(R1Aflag= ifelse(lotsize_sf>=7500,1,0),
         R1Bflag=ifelse(lotsize_sf<7500 & lotsize_sf>= 5000,1, 0)) %>% 
  count(R1Aflag, R1Bflag, zoning_code) %>% 
  knitr::kable()

```
Reassign some of the R1s to R1A and R1B based on the flag
```{r reassign R1}

vacantcleaned2 <- vacantcleaned %>% 
  mutate(R1Aflag= ifelse(lotsize_sf>=7500 & zoning_code=="R1",1,0),
         R1Bflag=ifelse(lotsize_sf<7500 & lotsize_sf>= 5000 & zoning_code=="R1",1, 0)) %>% 
  mutate(zoning_code= ifelse(R1Aflag==1, "R1A", zoning_code),
         zoning_code= ifelse(R1Bflag==1, "R1B", zoning_code))

#after cleaning there are still 1537 R1 that are not joined due to lot size not meeting standards. Not sure if we should keep some of them
anti_join(vacantcleaned2, zoningcode_factor, by = c("zoning_code")) %>% 
  filter(zoning_code=="R1") %>% 
  count(zoning_code) %>% 
  knitr::kable()

zoningdatanew2 <- left_join(vacantcleaned2, zoningcode_factor, by = "zoning_code")

zoningdatanew2 %>% 
  group_by(newdwelling_type) %>% 
  count() %>% 
  knitr::kable()
```

Filter out lots that are too small to build on and lots with missing zoning information, and select most dense zoning code for those that are big enough.Notice after filtering out lotsize_sf>= lot_area_min, observation dropped from 27520 to 12459, after slicing out the least dense zoning code, we only have 2131 observation left. This need more work. 

```{r filter-small}

zoningdata1 <- zoningdatanew2 %>% 
  mutate(lotsize_sf=(ifelse(is.na(lotsize_sf)==1,0, lotsize_sf)),
         lot_area_min=(ifelse(is.na(lot_area_min)==1,0, lot_area_min))) %>% 
  filter(!is.na(zoning_code),
         lotsize_sf >= lot_area_min) %>% 
  group_by(propaddress) %>% 
  slice(which.min(newdwelling_type)) %>% 
  ungroup()


# 13594 cases where lotsize_sf is less than the minimum lot area or missing. 
# by looking at the gap, should we come up with a threshold for a acceptable negative difference?
zoningdata1test <- zoningdatanew2 %>% 
    mutate(lotsize_sf=(ifelse(is.na(lotsize_sf)==1,0, lotsize_sf)),
         lot_area_min=(ifelse(is.na(lot_area_min)==1,0, lot_area_min))) %>% 
     filter(lotsize_sf < lot_area_min) %>% 
     mutate(gap= lotsize_sf-lot_area_min) %>% 
    select(lotsize_sf,lot_area_min, gap)


ggplot(zoningdata1, aes(lotsize_sf) )+
  geom_histogram(binwidth = 5000)


ggplot(zoningdata1, aes(lot_area_min) )+
  geom_histogram(binwidth = 500)


ggplot(zoningdata1test, aes(gap ))+
  geom_histogram(binwidth = 500)


```

## Calculate missing units

There are several variables we will use to calculate how many units we can put on each vacant lot.

1. Buildable area: this is `lotsize_sf` x the coverage ratio (`lot_coverage_max`)
2. Living area: this is the buildable area x the construction efficiency. The lower bound estimate is 70%, the upper bound estimate is 85%.
3. Units per floor: this is the living area divided by the average unit size. We use 900 sq. feet, based on the American Housing Survey 2017 data for the DC metro area.
4. Units allowed based on height limits: This is the units per floor x the number of stories.
5. Units allowed based on lot size: this is the lot area divided by `lotarea_per_unit`

We can create a function that will allow us to easily adjust these parameters and compare results.

```{r get-est}

#' Get maximum unit variables
#' @description this function generates the variables necessary to calculate the maximum number of units that can occupy one lot, per the by-right zoning code.
#'
#' @param dataset dataset with lot size and zoning variables 
#' @param construction_efficiency Assumption for construction efficiency
#' @param average_unit_size Assuption for average unit size- should be derived from the American Housing Survey. Given in square feet.
#'
#' @return returns a dataframe with new variables: `parameters`, `buildable_area`, `living_area`, `units_per_floor`, `floors`, `units_height`, `units_lotsize`, and `units_total`
#'
#' @examples
get_unit_variables <- function(dataset, 
                               construction_efficiency, 
                               average_unit_size) {
  
  dataset %>% 
    mutate(parameters = paste0("CE: ", construction_efficiency * 100,
                               "% , Unit size: ", average_unit_size,
                               " sf"),
           buildable_area = lotsize_sf * lot_coverage_max,
           living_area = buildable_area * construction_efficiency,
           units_per_floor = floor(living_area / average_unit_size),
           floors = floor(height_ft / 10),
           units_height = units_per_floor * floors,
           units_FAR= (living_area*floor_area_ratio_max)/average_unit_size,
          #lotarea_per_unit is missing for all in DC, so use units_height and units_FAR for units estimates
           units_lotsize = floor(lotsize_sf / lotarea_per_unit)) %>% 
    mutate_at(vars(units_height, units_FAR),
              funs(case_when(
                one_unit_flag == 1 ~ 1,
                two_unit_flag == 1 ~ 2,
                three_unit_flag == 1 ~ 3,
                four_unit_flag == 1 ~ 4,
                TRUE ~ .))) %>% 
    mutate(units_total = case_when(
      one_unit_flag == 1 ~ 1,
      two_unit_flag == 1 ~ 2,
      three_unit_flag == 1 ~ 3,
      four_unit_flag == 1 ~ 4,
      units_height > units_FAR ~ units_FAR,
      TRUE ~ units_height))
 
}

#' Get maximum unit summary
#'
#' @param dataset dataset with lot size and zoning variables 
#' @param construction_efficiency Assumption for construction efficiency
#' @param average_unit_size Assuption for average unit size- should be derived from the American Housing Survey. Given in square feet.
#' @param dwelling_type_breakdown Options are TRUE or FALSE- indicates whether or not estimates should be broken down by `dwelling_type`
#'
#' @return Returns a dataframe with summary statistics on the maximum units that can be build on lots, per the by-right zoning code.
#'
#' @examples
get_unit_summary <- function(dataset, 
                             construction_efficiency, 
                             average_unit_size,
                             dwelling_type_breakdown) {

  x <- dataset %>% 
    get_unit_variables(construction_efficiency,
                       average_unit_size) %>%
    group_by(parameters) %>% 
    summarize(lots = n(),
              units_height = sum(units_height),
              units_FAR = sum(units_FAR),
              units_lotsize = sum(units_lotsize),
              units_total = sum(units_total),
              units_per_lot = floor(sum(units_total / lots)))
  
  y <- dataset %>% 
    get_unit_variables(construction_efficiency,
                       average_unit_size) %>%
    group_by(parameters, dwelling_type) %>% 
    summarize(lots = n(),
              units_height = sum(units_height),
              units_FAR = sum(units_FAR),
              units_lotsize = sum(units_lotsize),
              units_total = sum(units_total),
              units_per_lot = floor(sum(units_total / lots))) %>% 
    mutate(dwelling_type = as.character(dwelling_type))
  
  
  if (dwelling_type_breakdown == TRUE) {
   
      bind_rows(y, mutate(x, dwelling_type = "All"))  
  
  } else if (dwelling_type_breakdown == FALSE) {
      x
    }
}



```

Get summary statistics for several different parameters

```{r sum}

get_unit_summary(dataset = zoningdata1,
                 construction_efficiency = .7,
                 average_unit_size = 900,
                 dwelling_type_breakdown = TRUE) %>% 
  knitr::kable()


map_df(c(.7, .85), ~ get_unit_summary(dataset = zoningdata1,
                                      construction_efficiency = .,
                                      average_unit_size = 900,
                                      dwelling_type_breakdown = TRUE)) %>% 
  knitr::kable()

```
Look at the distribution of units for lots with more than 1 unit, to check for extreme outliers.

```{r scatter, echo = FALSE}

zoningdata1 %>% 
  get_unit_variables(construction_efficiency = .7,
                     average_unit_size = 900) %>% 
             select(zoning_code, lotsize_sf, units_total, one_unit_flag, two_unit_flag, parameters, units_lotsize)%>% 
  filter(units_total > 1) %>% 
  ggplot() +
  geom_point(mapping = aes(lotsize_sf, units_total),
             alpha = .5) +
  scale_x_continuous(labels = scales::comma,
                     expand = expand_scale(mult = c(0, 0.002)), 
                     limits = c(0, 200000),
                     breaks = 0:6 * 50000) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.002)), 
                     limits = c(0, 300),
                     breaks = 0:7 * 50) +
  scatter_grid() +
  labs(x = "Lot size (square feet)", y = "Total units") +
  theme(plot.margin = margin(t = 20, r = 20))


zoningdata1test <- zoningdata1 %>% 
  get_unit_variables(construction_efficiency = .7,
                     average_unit_size = 900) %>% 
             select(zoning_code, lotsize_sf, units_total, one_unit_flag, two_unit_flag, parameters, units_lotsize)%>% 
  filter(units_total > 1) %>% 
  filter(two_unit_flag==1 & lotsize_sf>150000) 
# some of the higher lotsize_sf with only 2 units is due to the #dwelling units allowed flag. The three showed here are all RF codes that has a 2 dwelling units limit. I have yet to confirm with Chris that those are true. 


```

Look more closely at vacant lots < 50,000 SF

```{r scatter-small, echo = FALSE}

zoningdata1 %>% 
  get_unit_variables(construction_efficiency = .7,
                     average_unit_size = 900) %>% 
  filter(lotsize_sf < 50000) %>% 
  select(lotsize_sf, newdwelling_type, units_total)%>% 
  ggplot() +
  geom_point(mapping = aes(lotsize_sf, units_total,
                           color = newdwelling_type),
             alpha = .8) +
  scale_x_continuous(labels = scales::comma,
                     expand = expand_scale(mult = c(0, 0.0002)), 
                     limits = c(0, 50000),
                     breaks = 0:5 * 10000) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.0002)), 
                     limits = c(0, 60),
                     breaks = 0:4 * 15) +
  scatter_grid() +
  labs(x = "Lot size (square feet)", y = "Total units") +
  theme(plot.margin = margin(t = 20,
                             r = 20))

# agian the RF2 category has a two unit flag. Need to confirm with Chris. 


```

## Write out results

Write out the results to a csv. Include jurisdiction name in the title.

```{r write-results}

get_unit_summary(dataset = zoningdata1,
                 construction_efficiency = .7,
                 average_unit_size = 940,
                 dwelling_type_breakdown = TRUE) %>% 
  write_csv(paste0("L:/Libraries/RegHsg/Prog/Tables/",
                   filepath,
                   "-vacant-lots-analysis.csv"))


```