---
title: "Regional Housing Framework"
subtitle: "Vacant lots analysis for DC"
author: "YS"
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Yipeng adapted from Sarah Strochak

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The goal of this analysis is to determine how many units could be added by filling vacant lots with the most-dense developments allowed under the current zoning codes.
The two steps are:

1. Merge cleaned vacant lots data with zoning code data.
2. Calculate how many units are allowed one each vacant lot.

## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)
library(urbnthemes)
library(sf)

set_urbn_defaults("print")

source("../Macros/read-arlington.R")

```

Create directory for data exports
```{r dir}
if (!dir.exists("../Data")) {
  dir.create("../Data")
}

if (!dir.exists("L:/Libraries/RegHsg/Prog/Tables")) {
  dir.create("L:/Libraries/RegHsg/Prog/Tables")
}
```

Set FIPS code and filepath name
```{r fips}

currentfips <- "11001"
filepath <- "DC"
```


Load in cleaned Black Knight data for `r str_to_title(filepath)` County.

```{r read}

read_dc <- function(rmd = TRUE) {
  
  filename <- paste0("L:/Libraries/RegHsg/Data/", filepath,
                     "/", filepath, "-cleaned-data.csv")
  
  if (!file.exists(filename)) {
    stop("cleaned data not found in Data directory")
  } else {
    
    read_csv(filename,
             col_types = cols(county_fips = col_character(),
                              county_name = col_character(),
                              parcelbase_SSL = col_character(),
                              propaddress = col_character(),
                              propcity = col_character(),
                              propstate = col_character(),
                              propzip = col_character(),
                              propunitno = col_character(),
                              prophouseno = col_character(),
                              propstreetname = col_character(),
                              propstreetsuffix = col_character(),
                              lat = col_double(),
                              long = col_double(),
                              tract = col_character(),
                              owneroccupiedresidential = col_character(),
                              countylandusedescription = col_character(),
                              zoning = col_character(),
                              buildingarea = col_double(),
                              noofbuildings = col_character(),
                              noofstories = col_character(),
                              numberofunits = col_integer(),
                              yearbuilt = col_integer(),
                              lotsize_acres = col_double(),
                              lotsize_sf = col_double(),
                              address_type = col_character(),
                              category = col_character(),
                              category_detail = col_character(),
                              residential = col_integer(),
                              building_type = col_character(),
                              parcelbase_address = col_character(),
                              parcelbase_usecode = col_character(),
                              parcelgeo_x = col_double(),
                              parcelgeo_y = col_double(),
                              vacant_flag = col_integer()))
  }
}


if (!exists("jur")) {
  jur <- read_dc()
} else {
  warning(filepath, " data already read in")
}


```

## Clean zoning code information

### Move file 

First- **manually** move file from Box to the `zoning-codes` directory in the `Doc` folder on the L drive. Name file accordingly:
"jursidiction-zoning-code.csv"

### Read file

Read in zoning code, rename variables so that they are machine readable

```{r file-check}

if (!file.exists(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              " zoning code.csv"))) {
  
  stop("Move zoning file for jurisdiction to L:/Libraries/RegHsg/Doc/zoning-codes directory")
}

```


```{r read-zoning-csv}
z <- cols(
   `Code` = col_character(),
  `District Type` = col_character(),
  `Type of Dwelling` = col_character(),
  `Number of Dwelling units allowed` = col_number(),
  `Site Area` = col_number(),
  `Lot area` = col_number(),
  `Lot area per dwelling unit (sq. ft.) (minimum)` = col_number(),
  `Lot Width (average minimum, ft)` = col_character(),
  `Height (ft)` = col_double(),
  `Height (stories)` = col_double(),
  `Lot coverage maximum` = col_number(),
  `Floor area ratio max`=col_double(),
  `Maximum Density of Floor-Area-Ratio (FAR)` = col_character(),
  `Setbacks (centerline)` = col_character(),
  `Setbacks (right of way line)` = col_double(),
  `Side yard (ft)` = col_number(),
  `Side yard (both sides: minimum percent of required width)` = col_number(),
  `Rear yard` = col_double(),
  `Frontage` = col_double(),
  `Comments` = col_character()
)


zoningcode <- read_csv(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              " zoning code.csv"),
                       col_types = z)

names(zoningcode) <- 
  c("zoning_code", "district_type", "dwelling_type","dwelling_units_allowed",
    "site_area", "lot_area_min", "lotarea_per_unit",
    "lot_width", "height_ft", "height_stories",
    "lot_coverage_max","floor_area_ratio_max", "far_max",
    "setbacks_center", "setbacks_row",
    "front_side_yard", "side_yard_bothside","rear_side_yard", "frontage", "comments")

# fix variables read in as whole numbers (should be percent)
zoningcode <- zoningcode %>% 
  mutate_at(vars(lot_coverage_max, side_yard_bothside),
            ~ . / 100)
rm(z)

```

### Creating dwelling_type hierarchy

For the vacant lots analysis, we want to determine the `dwelling_type` that allows for the most dense construction for that zone. 

First, we look at all the options, these categories is not as helpful because attached and detached can mean so many different things plus we have a lot of NAs for residential code. So need to create a new dwelling type variable.

```{r tab-dwellings}
zoningcode %>% 
  count(dwelling_type) %>% 
  knitr::kable()
```

Count the number of unique zoning codes to make sure we won't get rid of any codes entirely.

```{r count-zoning}
length(unique(zoningcode$zoning_code))

```


The DC dwelling type is missing for a lot of zoning, so it is difficult to follow the exact same method. I will here create a new dwelling type variable based on some other variables in there. 

Since the variables are not always comparable, after the recoding, from most to least dense, order applies:

1. MF
2. RF3 (Residential Flat4)
3. RF2 (Residential Flat2)
4. RF1 (Residential Flat1)
5. One Family 

```{r recode dwelling}
zoningcode1 <- zoningcode %>% 
  mutate(newdwelling_type = case_when(
    dwelling_units_allowed==1 ~ "One Family",
    dwelling_units_allowed==2 ~ "RF1",
    dwelling_units_allowed==3 ~ "RF2",
    dwelling_units_allowed==4 ~ "RF3",
    floor_area_ratio_max>1 & is.na(dwelling_units_allowed) ~ "MF",
    is.na(floor_area_ratio_max) &is.na(dwelling_units_allowed) ~ "All Other",
    floor_area_ratio_max <= 1 ~ "Other Mixed",
    zoning_code=="RA-1"|zoning_code=="RA-6" ~ "RF1"
  ))

```
Filter out dwelling_types of "All", "All Other", and those that are dependent on a shared boundary.

```{r purge}
zoningcode2 <- zoningcode1 %>% 
  group_by(zoning_code) %>% 
  filter(!(newdwelling_type %in% c("All Other",
                               "Other Mixed")
           & n() > 1)) %>% 
  ungroup()

```

 Make sure we did not lose any zoning codes entirely.
```{r responsible-purge}

stopifnot(length(unique(zoningcode$zoning_code))
          == length(unique(zoningcode1$zoning_code)))

```

### Create factor variable
To use this hierarchy, we transform the `newdwelling_type` variable into a factor, assign the factors to the appropriate level, (1-6, as listed above).
We will also create a designation for dwelling types that are restricted to a certain amount of units (one unit for "One Family Dwelling" and one for two units for "Duplex"). 

**NOTE** This will be VERY jurisdiction specific.

**NOTE** If there are no one or two unit limitations, still make the variable and set them all equal to zero. This allows the functionality of the calculation to be more uniform across counties.

```{r dwelling-factor}

zoningcode1 <- zoningcode1 %>% 
  mutate( newdwelling_type = factor(newdwelling_type,
                                levels = c("MF",
                                           "RF3",
                                           "RF2",
                                           "RF1",
                                           "One Family")),
         one_unit_flag = ifelse(newdwelling_type == "One Family",
                                1,
                                0),
         two_unit_flag = ifelse(newdwelling_type == "RF2"|newdwelling_type == "RF3",
                                1,
                                0))

```
