---
title: "Regional Housing Framework"
subtitle: "Vacant lots analysis for Montgomery County"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Patrick Spauster
 
Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The goal of this analysis is to determine how many units could be added by filling vacant lots with the most-dense developments allowed under the current zoning codes.
The two steps are:

1. Merge cleaned vacant lots data with zoning code data.
2. Calculate how many units are allowed one each vacant lot.

## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)
library(urbnthemes)
library(sf)
library(ggplot2)
library(urbnthemes)

set_urbn_defaults("print")

source("../Macros/read-montgomery.R")

```

Create directory for data exports
```{r dir}
if (!dir.exists("../Data")) {
  dir.create("../Data")
}

if (!dir.exists("L:/Libraries/RegHsg/Prog/Tables")) {
  dir.create("L:/Libraries/RegHsg/Prog/Tables")
}
```

Set FIPS code and filepath name
```{r fips}

currentfips <- "24031"
filepath <- "montgomery"


```

Load in cleaned Black Knight data for `r str_to_title(filepath)` County.

```{r read}
if (!exists("jur")) {
  jur <- read_montgomery()
} else {
  warning(filepath, " data already read in")
}

```


## Clean zoning code information

### Move file 

First- **manually** move file from Box to the `zoning-codes` directory in the `Doc` folder on the L drive. Name file accordingly:
"jursidiction-zoning-code.csv"

### Read file

Read in zoning code, rename variables so that they are machine readable

```{r file-check}

if (!file.exists(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              "-zoning-code.csv"))) {
  
  stop("Move zoning file for jurisdiction to L:/Libraries/RegHsg/Doc/zoning-codes directory")
}

```

```{r read-zoning-csv}
z <- cols(
  Jur = col_character(),
  Code = col_character(),
  `Zone` = col_character(),
  `Type of Dwelling` = col_character(),
  `Site Area` = col_number(),
  `Site area per dwelling unit (sq. ft.) (minimum)` = col_number(),
  `Lot area` = col_number(),
  `Lot area per dwelling unit (sq. ft.) (minimum)` = col_number(),
  `Lot Width (average minimum, ft)` = col_character(),
  `Lot Width building line (average minimum, ft)` = col_character(),
  `Height (ft)` = col_double(),
  `Site coverage maximum` = col_number(),
  `Lot coverage maximum` = col_number(),
  `Floor area minimum` = col_double(),
  `Maximum Density of Floor-Area-Ratio (FAR)` = col_character(),
  `Setbacks (front)` = col_character(),
  `Side (ft)` = col_character(),
  `Rear (ft)` = col_character(),
  `Density (units/acre)` = col_number(),
  `Massing Units/Row` = col_number(),
  `Open Space minimum percentage` = col_number(),
  Comments = col_character()
)

zoningcode <- read_csv(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              "-zoning-code.csv"),
                       col_types = z)

names(zoningcode) <- 
  c("juristdiction", "zoning_code", "zone", "dwelling_type",
    "site_area_min", "sitearea_per_unit", "lot_area_min", "lotarea_per_unit",
    "lot_width","lot_width_building", "height_ft",
    "site_coverage_max", "lot_coverage_max", "floor_area_min", "far_max",
    "setbacks_front", "setbacks_side",
    "setbacks_rear", "density", "massing", "open_space", "comments")

# fix variables read in as whole numbers (should be percent)
zoningcode <- zoningcode %>% 
  mutate_at(vars(lot_coverage_max, site_coverage_max, open_space),
            ~ . / 100)
rm(z)

```

### Creating dwelling_type hierarchy

For the vacant lots analysis, we want to determine the `dwelling_type` that allows for the most dense construction for that zone. 

First, we look at all the options, and filter out all the observations marked as "other", since we cannot be sure what property type that refers to. We will also filter out categories that are dependent on adjacent lot lines.

```{r tab-dwellings}

zoningcode %>% 
  count(dwelling_type, juristdiction) %>% 
  knitr::kable()
```

Count the number of unique zoning codes to make sure we won't get rid of any codes entirely.

```{r count-zoning}
length(unique(zoningcode$zoning_code))

```

```{r zoning by juristdiction}

zoning_montgomery <- zoningcode %>% 
  filter(juristdiction == "Montgomery")

zoning_rockville <- zoningcode %>% 
  filter(juristdiction == "Rockville")

zoning_gaithersburg <- zoningcode %>% 
  filter(juristdiction == "Gaithersburg")


```




*For Montgomery*
```{r list codes}
zoning_montgomery %>% 
  count(dwelling_type) %>% 
  knitr::kable()

```


Since the variables are not always comparable, the following order, from most to least dense, applies:

1. apartment
2. townhouse
3. duplex-over
4. duplex-side
5. detached house

multi-use and general have the same zoning specs as apartment. I change all of the multi-use and general zoning codes to apartments for simplicity.

```{r edit mont codes}

zoning_montgomery <- zoning_montgomery %>% 
  mutate(dwelling_type = ifelse(dwelling_type == "general"| dwelling_type == "multi-use", "apartment", dwelling_type))

```
Check to see if they were replaced
```{r list codes}
zoning_montgomery %>% 
  count(dwelling_type) %>% 
  knitr::kable()

```


### Create factor variable
To use this hierarchy, we transform the `dwelling_type` variable into a factor, assign the factors to the appropriate level, (1-6, as listed above).
We will also create a designation for dwelling types that are restricted to a certain amount of units (one unit for "One Family Dwelling" and one for two units for "Duplex"). 

**NOTE** This will be VERY jurisdiction specific.

**NOTE** If there are no one or two unit limitations, still make the variable and set them all equal to zero. This allows the functionality of the calculation to be more uniform across counties.

```{r dwelling-factor}

zoning_montgomery1 <- zoning_montgomery %>% 
  mutate(dwelling_type=ifelse(is.na(dwelling_type)==1, "all", dwelling_type)) %>% 
  mutate(dwelling_type = factor(dwelling_type,
                                levels = c("apartment",
                                           "townhouse",
                                           "duplex-over",
                                           "duplex-side",
                                           "detached house"
                                           )),
         one_unit_flag = ifelse(dwelling_type == "detached house",
                                1,
                                0),
         two_unit_flag = ifelse(dwelling_type == "duplex-over" | dwelling_type == "duplex-side",
                                1,
                                0))

```

*For Gaithersburg*
```{r list codes}
zoning_gaithersburg %>% 
  count(dwelling_type) %>% 
  knitr::kable()

```


Since the variables are not always comparable, the following order, from most to least dense, applies:

1. multiple family condominium dwellings
2. multiple family dwelling 
3. townhouse
4. two family dwellings
5. single-family attached 
6. single-family semidetached 
7. single-family detached 
8. all (dwelling type is NA)

### Create factor variable
To use this hierarchy, we transform the `dwelling_type` variable into a factor, assign the factors to the appropriate level, (1-6, as listed above).
We will also create a designation for dwelling types that are restricted to a certain amount of units (one unit for "One Family Dwelling" and one for two units for "Duplex"). 

**NOTE** This will be VERY jurisdiction specific.

**NOTE** If there are no one or two unit limitations, still make the variable and set them all equal to zero. This allows the functionality of the calculation to be more uniform across counties.

```{r dwelling-factor}

zoning_gaithersburg1 <- zoning_gaithersburg %>% 
  mutate(dwelling_type = factor(dwelling_type,
                                levels = c("multiple family condominium dwellings",
                                           "multiple family dwelling",
                                           "townhouse",
                                           "two family dwellings",
                                           "single-family attached",
                                           "single-family semidetached",
                                           "single-family detached",
                                           "all"
                                           )),
         one_unit_flag = ifelse(dwelling_type == "single-family attached" | 
                                dwelling_type == "single-family semidetached" | 
                                dwelling_type == "single-family detached",
                                1,
                                0),
         two_unit_flag = ifelse(dwelling_type == "two family dwellings",
                                1,
                                0))


```

*For Rockville*
```{r list codes}
zoning_rockville %>% 
  count(dwelling_type) %>% 
  knitr::kable()

```


There are no dwelling types, so the factor is for all.

### Create factor variable
All rockville zoning codes are unique, so we do not need a factor. We will automatically select "all", or treat it as dwelling type is NA

**NOTE** This will be VERY jurisdiction specific.

**NOTE** If there are no one or two unit limitations, still make the variable and set them all equal to zero. This allows the functionality of the calculation to be more uniform across counties.

```{r dwelling-factor}

zoning_rockville1 <- zoning_rockville %>% 
  mutate(dwelling_type = factor(dwelling_type,
                                levels = c(
                                           "all"
                                           )))


```







## Clean vacant lots data

Select only vacant lots

```{r res}

vacant <- jur %>% 
  filter(category == "vacant", propcity != "GAITHERSBURG", propcity != "ROCKVILLE", propcity != "ROCVKILE", propcity != "GAITHERBURG")

x <- count(vacant, propcity)
```


See how many lots have missing key variables

```{r missing}

vacant %>% 
  mutate(num = NA) %>% 
  group_by(category_detail) %>% 
  summarize_at(vars(propaddress, lotsize_sf, zoning, num),
               ~ sum(is.na(.))) %>% 
  mutate_at(vars(propaddress, lotsize_sf, zoning),
            .funs = funs(percent = scales::percent(. / num))) %>% 
  knitr::kable(caption = "Number and percent of missing values")

```

### Assessed Value
I removed the assessed value section because there were no relevant properties with structures that needed to be removed fromt he list.



## Merge with cleaned data

Now that we have selected the correct records from the zoning code file, we can merge with the cleaned property records data. First, we have to do a bit of cleaning of these codes, including getting rid of instances where there is a tie, and correcting some of the notation. We will create a new variable in order to archive the original `zoning` variable.

```{r bk-clean}
mvacant <- vacant1 %>% 
  filter()
z <- count(vacant1, zoning)



vacantfinal <- vacant1 %>% 
  mutate(zoning_code = case_when(
    zoning == "CR1.0" ~ "CR",
    zoning %in% c("CRT0.5","CRT075", "CRT1.5") ~ "CRT",
    zoning %in% c("R3", "R30") ~ "R-30",
    zoning %in% c("R200T", "R200") ~ "R-200",
    zoning %in% c("R20") ~ "R-20",
    zoning %in% c("R40") ~ "R-40",
    zoning %in% c("R90","R90TD") ~ "R-90",
    zoning %in% c("R60") ~ "R-60",
    zoning %in% c("R10") ~ "R-10",
    zoning %in% c("RE2") ~ "Re-2",
    zoning %in% c("RE2C") ~ "Re-2C",
    zoning %in% c("RE1") ~ "Re-1",
    zoning %in% c("RURAL", "RUR") ~ "R",
    TRUE ~ zoning
  ))

z <- count(vacantfinal, zoning_code)
           
```


Merge on zoning code variable. Test that all variables merged.

```{r zoning-merge}

zoningdata <- left_join(vacantfinal, zoning_montgomery1, by = "zoning_code")

count(zoningdata, zoning_code)

```

There are `r ` observations that did not have a match in the zoning code. Below, we see which zoning codes were in the Black Knight data but NOT in the zoning code data.

```{r antijoin}

anti_join(vacantfinal, zoning_montgomery1, by = c("zoning_code")) %>% 
  count(zoning) %>% 
  knitr::kable()

```

Aside from those with missing zoning information, these zoning codes do not allow for residential properties, which explains why they were not found in the zoning code data.

Filter out lots that are too small to build on and lots with missing zoning information, and select most dense zoning code for those that are big enough.

```{r filter-small}

zoningdata1 <- zoningdata %>% 
  filter(!is.na(zoning),
         lotsize_sf >= lot_area_min) %>% 
  group_by(propaddress) %>% 
  slice(which.min(dwelling_type)) %>% 
  ungroup()

```


## Calculate missing units

before we do the calculation, the montgomery density measure uses lotsize/acre. we need to fill in the lotsize_acre variable for those lot sizes that we filled in from the parcel file.
```{r acres}
zoningdata2 <- zoningdata1 %>% 
  mutate(lotsize_acres = if_else(is.na(lotsize_acres), lotsize_sf / 43560, lotsize_acres)) %>% 
  mutate(lot_area_min = if_else(is.na(lot_area_min), site_area_min, lot_area_min)) %>% 
  mutate(lotarea_per_unit = if_else(is.na(lotarea_per_unit), sitearea_per_unit, lotarea_per_unit)) %>% 
  mutate(lot_coverage_max = if_else(is.na(lot_coverage_max), site_coverage_max, lot_coverage_max))


```


There are several variables we will use to calculate how many units we can put on each vacant lot.

1. Buildable area: this is `lotsize_sf` x the coverage ratio (`lot_coverage_max`)
2. Living area: this is the buildable area x the construction efficiency. The lower bound estimate is 70%, the upper bound estimate is 85%.
3. Units per floor: this is the living area divided by the average unit size. We use 900 sq. feet, based on the American Housing Survey 2017 data for the DC metro area.
4. Units allowed based on height limits: This is the units per floor x the number of stories.
3. Units allowed based on lot size: this is the lot area divided by `lotarea_per_unit`

We can create a function that will allow us to easily adjust these parameters and compare results.

```{r get-est}

#' Get maximum unit variables
#' @description this function generates the variables necessary to calculate the maximum number of units that can occupy one lot, per the by-right zoning code.
#'
#' @param dataset dataset with lot size and zoning variables 
#' @param construction_efficiency Assumption for construction efficiency
#' @param average_unit_size Assuption for average unit size- should be derived from the American Housing Survey. Given in square feet.
#'
#' @return returns a dataframe with new variables: `parameters`, `buildable_area`, `living_area`, `units_per_floor`, `floors`, `units_height`, `units_lotsize`, and `units_total`
#'
#' @examples
get_unit_variables <- function(dataset, 
                               construction_efficiency, 
                               average_unit_size) {
  
  dataset %>% 
    mutate(parameters = paste0("CE: ", construction_efficiency * 100,
                               "% , Unit size: ", average_unit_size,
                               " sf"),
           buildable_area = lotsize_sf * lot_coverage_max,
           living_area = buildable_area * construction_efficiency,
           units_per_floor = floor(living_area / average_unit_size),
           floors = floor(height_ft / 10),
           units_height = units_per_floor * floors,
           units_density = floor(density * lotsize_acres),
           units_lotsize = floor(lotsize_sf / lotarea_per_unit)) %>% 
    mutate_at(vars(units_height, units_lotsize),
              funs(case_when(
                one_unit_flag == 1 ~ 1,
                two_unit_flag == 1 ~ 2,
                TRUE ~ .))) %>% 
    mutate(units_total = case_when(
      one_unit_flag == 1 ~ 1,
      two_unit_flag == 1 ~ 2,
      TRUE ~ pmin(units_lotsize, units_height, units_density)))
  
}

#' Get maximum unit summary
#'
#' @param dataset dataset with lot size and zoning variables 
#' @param construction_efficiency Assumption for construction efficiency
#' @param average_unit_size Assuption for average unit size- should be derived from the American Housing Survey. Given in square feet.
#' @param dwelling_type_breakdown Options are TRUE or FALSE- indicates whether or not estimates should be broken down by `dwelling_type`
#'
#' @return Returns a dataframe with summary statistics on the maximum units that can be build on lots, per the by-right zoning code.
#'
#' @examples
get_unit_summary <- function(dataset, 
                             construction_efficiency, 
                             average_unit_size,
                             dwelling_type_breakdown) {

  x <- dataset %>% 
    get_unit_variables(construction_efficiency,
                       average_unit_size) %>%
    group_by(parameters) %>% 
    summarize(lots = n(),
              units_height = sum(units_height, na.rm = TRUE),
              units_lotsize = sum(units_lotsize, na.rm = TRUE),
              units_total = sum(units_total, na.rm = TRUE),
              units_per_lot = floor(sum(units_total / lots, na.rm = TRUE)))
  
  y <- dataset %>% 
    get_unit_variables(construction_efficiency,
                       average_unit_size) %>%
    group_by(parameters, dwelling_type) %>% 
    summarize(lots = n(),
              units_height = sum(units_height, na.rm = TRUE),
              units_lotsize = sum(units_lotsize, na.rm = TRUE),
              units_total = sum(units_total, na.rm = TRUE),
              units_per_lot = floor(sum(units_total / lots, na.rm = TRUE))) %>% 
    mutate(dwelling_type = as.character(dwelling_type))
  
  
  if (dwelling_type_breakdown == TRUE) {
   
      bind_rows(y, mutate(x, dwelling_type = "All"))  
  
  } else if (dwelling_type_breakdown == FALSE) {
      x
    }
}



```

```{r sum}
results <- get_unit_variables(dataset = zoningdata2,
                            construction_efficiency = .7,
                            average_unit_size = 940)


#results %>% 
#  mutate(num = NA) %>% 
#  group_by(category_detail) %>% 
#  summarize_at(vars(lot_area_min, lotarea_per_unit, zoning, num),
#               ~ sum(is.na(.))) %>% 
#  mutate_at(vars(propaddress, lotsize_sf, zoning),
#            .funs = funs(percent = scales::percent(. / num))) %>% 
#  knitr::kable(caption = "Number and percent of missing values")
```
 

Get summary statistics for several different parameters

```{r sum}

get_unit_summary(dataset = zoningdata2,
                 construction_efficiency = .7,
                 average_unit_size = 940,
                 dwelling_type_breakdown = TRUE) %>% 
  knitr::kable()


map_df(c(.7, .85), ~ get_unit_summary(dataset = zoningdata2,
                                      construction_efficiency = .,
                                      average_unit_size = 940,
                                      dwelling_type_breakdown = FALSE)) %>% 
  knitr::kable()

```

Look at the distribution of units for lots with more than 1 unit, to check for extreme outliers.

```{r scatter, echo = FALSE}

zoningdata2 %>% 
  get_unit_variables(construction_efficiency = .7,
                     average_unit_size = 940) %>% 
  filter(units_total > 1) %>% 
  ggplot() +
  geom_point(mapping = aes(lotsize_sf, units_total),
             alpha = .5) +
  scale_x_continuous(labels = scales::comma,
                     expand = expand_scale(mult = c(0, 0.002)), 
                     limits = c(0, 300000),
                     breaks = 0:6 * 50000) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.002)), 
                     limits = c(0, 100),
                     breaks = 0:7 * 50) +
  #scatter_grid() +
  labs(x = "Lot size (square feet)", y = "Total units") +
  theme(plot.margin = margin(t = 20, r = 20))

```


Look more closely at vacant lots < 50,000 SF

```{r scatter-small, echo = FALSE}

zoningdata2 %>% 
  get_unit_variables(construction_efficiency = .7,
                     average_unit_size = 940) %>% 
  filter(lotsize_sf < 50000,
         units_total > 1) %>% 
  ggplot() +
  geom_point(mapping = aes(lotsize_sf, units_total,
                           color = dwelling_type),
             alpha = .8) +
  scale_x_continuous(labels = scales::comma,
                     expand = expand_scale(mult = c(0, 0.0002)), 
                     limits = c(0, 50000),
                     breaks = 0:5 * 10000) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.0002)), 
                     limits = c(0, 60),
                     breaks = 0:4 * 15) +
  #scatter_grid() +
  labs(x = "Lot size (square feet)", y = "Total units") +
  theme(plot.margin = margin(t = 20,
                             r = 20))



```

## Write out results

Write out the results to a csv. Include jurisdiction name in the title.

```{r write-results}

get_unit_summary(dataset = zoningdata2,
                 construction_efficiency = .7,
                 average_unit_size = 940,
                 dwelling_type_breakdown = TRUE) %>% 
  write_csv(paste0("L:/Libraries/RegHsg/Prog/Tables/",
                   filepath,
                   "-vacant-lots-analysis.csv"))


```
