---
title: "Regional Housing Framework"
subtitle: "Vacant lots analysis for Arlington County"
author: ""
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: ../../www/web_report.css
    editor_options:
      chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

Library: RegHsg

Project: Regional Housing Framework

Author: Sarah Strochak

Version: R 3.5.1, RStudio 1.1.423

Last updated `r format(Sys.time(), '%B %d, %Y')`

Environment: Local Windows session (desktop)


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# knitr::opts_chunk$set(eval = FALSE)
```

## Description

The goal of this analysis is to determine how many units could be added by filling vacant lots with the most-dense developments allowed under the current zoning codes.
The two steps are:

1. Merge cleaned parcel data with zoning code data.
2. Calculate how many units are allowed one each parcel.
3. Filter out and sum for vacant lots.

## Set-up
Load libraries and functions
```{r setup}
library(tidyverse)
library(urbnthemes)
library(sf)

set_urbn_defaults("print")

source("../../Macros/read-arlington.R")

```

Create directory for data exports
```{r dir}
if (!dir.exists("../../Data")) {
  dir.create("../../Data")
}

if (!dir.exists("L:/Libraries/RegHsg/Prog/Tables")) {
  dir.create("L:/Libraries/RegHsg/Prog/Tables")
}
```

Set FIPS code and filepath name
```{r fips}

currentfips <- "51013"
filepath <- "arlington"


```

Load in cleaned Black Knight data for `r str_to_title(filepath)` County.

```{r read}
if (!exists("jur")) {
  jur <- read_arlington()
} else {
  warning(filepath, " data already read in")
}

```


## Clean zoning code information

### Move file 

First- **manually** move file from Box to the `zoning-codes` directory in the `Doc` folder on the L drive. Name file accordingly:
"jursidiction-zoning-code.csv"

### Read file

Read in zoning code, rename variables so that they are machine readable

```{r file-check}

if (!file.exists(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              "-zoning-code.csv"))) {
  
  stop("Move zoning file for jurisdiction to L:/Libraries/RegHsg/Doc/zoning-codes directory")
}

```

```{r read-zoning-csv}
z <- cols(
  Code = col_character(),
  `District Type` = col_character(),
  `Type of Dwelling` = col_character(),
  `Site Area` = col_number(),
  `Lot area` = col_number(),
  `Lot area per dwelling unit (sq. ft.) (minimum)` = col_number(),
  `Lot Width (average minimum, ft)` = col_character(),
  `Height (ft)` = col_double(),
  `Height (stories)` = col_double(),
  `Lot coverage maximum` = col_number(),
  `Floor area minimum` = col_double(),
  `Maximum Density of Floor-Area-Ratio (FAR)` = col_character(),
  `Setbacks (centerline)` = col_character(),
  `Setbacks (right of way line)` = col_double(),
  `Front/Side yard (ft)` = col_double(),
  `Side yard (both sides: minimum percent of required width)` = col_number(),
  Frontage = col_double(),
  Comments = col_character()
)

zoningcode <- read_csv(paste0("L:/Libraries/RegHsg/Doc/zoning-codes/",
                              filepath,
                              "-zoning-code.csv"),
                       col_types = z)

names(zoningcode) <- 
  c("zoning_code", "district_type", "dwelling_type",
    "site_area", "lot_area_min", "lotarea_per_unit",
    "lot_width", "height_ft", "height_stories",
    "lot_coverage_max", "floor_area_min", "far_max",
    "setbacks_center", "setbacks_row",
    "front_side_yard", "side_yard", "frontage", "comments")

# fix variables read in as whole numbers (should be percent)
zoningcode <- zoningcode %>% 
  mutate_at(vars(lot_coverage_max, side_yard),
            ~ . / 100)
rm(z)

```

### Creating dwelling_type hierarchy

For the vacant lots analysis, we want to determine the `dwelling_type` that allows for the most dense construction for that zone. 

First, we look at all the options, and filter out all the observations marked as "other", since we cannot be sure what property type that refers to. We will also filter out categories that are dependent on adjacent lot lines.

```{r tab-dwellings}

zoningcode %>% 
  count(dwelling_type) %>% 
  knitr::kable()

```

Count the number of unique zoning codes to make sure we won't get rid of any codes entirely.

```{r count-zoning}
length(unique(zoningcode$zoning_code))

```
Filter out dwelling_types of "All", "All Other", and those that are dependent on a shared boundary.

```{r purge}
zoningcode1 <- zoningcode %>% 
  group_by(zoning_code) %>% 
  filter(!(dwelling_type %in% c("All",
                                "All Other",
                                "Duplex; share a lot line with RA, C, or M",
                                "Semi-detatched; share a lot line with RA, C, or M")
           & n() > 1)) %>% 
  ungroup()

```

 Make sure we did not lose any zoning codes entirely.
```{r responsible-purge}

stopifnot(length(unique(zoningcode$zoning_code))
          == length(unique(zoningcode1$zoning_code)))

```

Since the variables are not always comparable, the following order, from most to least dense, applies:

1. Multiple-family
2. Townhouse
3. Semi-detached
4. Duplex
5. Other Residential
6. One Family Dwelling

### Create factor variable
To use this hierarchy, we transform the `dwelling_type` variable into a factor, assign the factors to the appropriate level, (1-6, as listed above).
We will also create a designation for dwelling types that are restricted to a certain amount of units (one unit for "One Family Dwelling" and one for two units for "Duplex"). 

**NOTE** This will be VERY jurisdiction specific.

**NOTE** If there are no one or two unit limitations, still make the variable and set them all equal to zero. This allows the functionality of the calculation to be more uniform across counties.

```{r dwelling-factor}

zoningcode1 <- zoningcode1 %>% 
  mutate(dwelling_type = factor(dwelling_type,
                                levels = c("Multiple-family",
                                           "Townhouse",
                                           "Semi-detached",
                                           "Duplex",
                                           "Other Residential",
                                           "One Family Dwelling")),
         one_unit_flag = ifelse(dwelling_type == "One Family Dwelling",
                                1,
                                0),
         two_unit_flag = ifelse(dwelling_type == "Duplex",
                                1,
                                0))

```


## Clean parcel data

See how many lots have missing key variables

```{r missing}

jur %>% 
  mutate(num = NA) %>% 
  group_by(category_detail) %>% 
  summarize_at(vars(propaddress, lotsize_sf, zoning, num),
               ~ sum(is.na(.))) %>% 
  mutate_at(vars(propaddress, lotsize_sf, zoning),
            .funs = funs(percent = scales::percent(. / num))) %>% 
  knitr::kable(caption = "Number and percent of missing values")

```

### Missing lot size
We will fill in the missing lot size, when we can, with the area of the lot, obtained from the parcel shapefile in the precleaning step. If that is also missing, we use the lot size variables from the property file. All variables are in square feet.

```{r fill-lots}

jur1 <- jur %>% 
  mutate(arc_lotsize_sf = lotsize_sf, # archive the original lot size variable
         lotsize_sf = ifelse(is.na(lotsize_sf),
                             parcel_area,
                             lotsize_sf),
         lotsize_sf = ifelse(is.na(lotsize_sf),
                             lotsize_prop,
                             lotsize_sf))

```

See how many were filled in

```{r count-lots}
sum(is.na(jur$lotsize_sf)) - sum(is.na(jur1$lotsize_sf))
```

### Assessment values

The difference between the land value, improvement value, and total value can help us determine if lots are really vacant.

```{r check-assess}

jur1 %>% 
  group_by(category_detail) %>% 
  summarize(nprops = n(),
            imp_zero = sum(assessedimprovementvalue == 0, na.rm = TRUE),
            imp_na = sum(is.na(assessedimprovementvalue))) %>% 
  mutate(imp_none = imp_zero + imp_na,
         percent_none = imp_none / nprops) %>% 
  knitr::kable()
  

```


Most of the vacant properties have either a 0 or a missing value for the improvement value, which is what we would expect. We can take a closer look at properties that have a non-zero value and decide if we need to re-categorize them as not really vacant- if there is a structure there, we do not want to include it in this calculation.

```{r not-so-vacant}

jur1 %>% 
  filter(category == "vacant",
         assessedimprovementvalue > 0, 
         !is.na(assessedimprovementvalue)) %>% 
  mutate(p_total = assessedimprovementvalue / totalassessedvalue) %>% 
  group_by(category_detail) %>% 
  summarize(n = n(),
            mean = mean(assessedimprovementvalue),
            median = median(assessedimprovementvalue),
            min = min(assessedimprovementvalue),
            max = max(assessedimprovementvalue),
            p_total = median(p_total)) %>% 
  knitr::kable()

```

Look at the distribution of improved value as a percent of total value.

```{r scatter-not-vacant, echo=FALSE}
jur1 %>% 
  filter(category == "vacant",
         assessedimprovementvalue > 0, 
         !is.na(assessedimprovementvalue)) %>% 
  mutate(p_total = assessedimprovementvalue / totalassessedvalue) %>% 
  ggplot() +
  geom_histogram(mapping = aes(p_total),
                 bins = 30) +
  scale_x_continuous(labels = scales::percent,
                     breaks = c(0:10 *.1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Percent of total assessed value that is derived from improvement value",
       y = NULL)
```

We can recategorize these observations as "vacant with structure", and will not include these in the calculation. We can also make a flag for missing coordinates.

```{r not-vacant-recat}

jur2 <- jur1 %>% 
  mutate(category = ifelse(assessedimprovementvalue > 0 & 
                           !is.na(assessedimprovementvalue) &
                             category == "vacant",
                           "vacant with structure",
                           category))

count(jur2, category)

jur2 <- jur2 %>% 
  mutate(missing_coord = ifelse(is.na(lat) | is.na(long),
                                1,
                                0))
```

### Missing zoning variables

To determine zoning codes for addresses that have a missing zoning designation or multiple zoning designations, we will use a spatial join to the county's zoning layer.

Read in zoning layer, set CRS
```{r read-zoning}
zdir <- paste0("L:/Libraries/RegHsg/Maps/",
               filepath,
               "-zoning-map")
zfile <- str_sub(list.files(zdir), end = -5) %>% unique()

zon <- read_sf(dsn = zdir,
               layer = zfile)

st_geometry(zon)
zon <- st_transform(zon, crs = 4326)

```

Convert parcel data to SF, set CRS. To convert to SF, we need to use only the observations with non-missing longitude and latitude. We separate these off and recombine after the spatial join.

```{r convert-sf}

jursf <- jur2 %>% 
  filter(missing_coord == 0) %>% 
  st_as_sf(coords = c("long", "lat")) %>% 
  st_set_crs(st_crs(zon))
```

Plot to make sure they are overlapping
```{r make-map, echo=FALSE}
ggplot() +
  geom_sf(data = zon, mapping = aes(fill = GZDC)) +
  scale_fill_manual(values = c("#1696d2", "#fdbf11", 
                               "#d2d2d2", "#ec008b", 
                               "#55b748")) +
  geom_sf(data = sample_n(jursf, 500), mapping = aes(),
          color = "black")

```

Spatial join

```{r spatial-join}

jurjoin <- st_join(jursf, zon, join = st_intersects)

```

After the join, we convert back to a dataframe, for ease of computations.
```{r back-to-df}

jur3 <- st_set_geometry(jurjoin, NULL)

```

Count how many observations did not join. This will throw an error if more than 5% did not match- in this case, check the underlying shapefiles for compatability.

```{r count-spatial}

paste0(sum(is.na(jur3$ZN_DESIG)), " observations didn't join- ",
      round(sum(is.na(jur3$ZN_DESIG)) / nrow(jur3) * 100, 2), "%")

if (sum(is.na(jur3$ZN_DESIG)) / nrow(jur3) > .05) {
  stop("More than 5% of observations could not be spatially joined.")
}
      
```

Test how many have different results from the Black Knight zoning variable.

```{r test-spatial}

jur3 %>% 
  group_by(is.na(zoning)) %>% 
  count(zoning == ZN_DESIG)

```

Take a closer look how many and what percent are different "just because"- not because they didn't merge, have missing values, or have duplicate values.

```{r test-spatial-again}

jur3 %>% 
  filter(!is.na(zoning), 
         !is.na(ZN_DESIG), 
         str_detect(zoning, "/") == FALSE) %>% 
  mutate(different_flag = ifelse(zoning != ZN_DESIG,
                                 1,
                                 0)) %>% 
  summarize(different = sum(different_flag, na.rm = TRUE)) %>% 
  mutate(percent_different = paste0(round(different / nrow(jur3) *100, 2), "%"))

```

Many are different because the Black Knight zoning has two zoning codes, separated by a slash. We can test how many of these instances have the spatially joined zoning designation in the Black Knight zoning variable.

```{r slashes}

jur3 %>% 
  filter(str_detect(zoning, "/")) %>% 
  count(str_detect(zoning, ZN_DESIG))

```

If the zoning is missing OR the zoning contains a slash and the spatially joined zoning designation is one of the dual descriptions, we replace with spatial joined zoning. Archive Black Knight zoning variable for future reference.

```{r zoning-swap}

jur4 <- jur3 %>% 
  mutate(zoning_bk = zoning,
         zoning = ifelse(is.na(zoning), 
                         ZN_DESIG, 
                         zoning),
         zoning = ifelse(str_detect(zoning, "/") == TRUE &
                           str_detect(zoning_bk, ZN_DESIG) == TRUE,
                         ZN_DESIG,
                         zoning)) %>% 
  select(-c(OBJECTID, ZONECODE, LABEL, SYMBOL, ZN_DESIG_6, GZDC))

```

Recombine with observations with missing coordinates, check that number of rows is the same.

```{r get-final}

jurtotal <- jur2 %>% 
  filter(is.na(lat) | is.na(long)) %>% 
  mutate(zoning_bk = zoning) %>% 
  bind_rows(jur4)

stopifnot(nrow(jurtotal) == nrow(jur))

```

### Dual zoning codes

The last step is to identify a zoning designation for observations that either...

1. Have missing coordinates and have a dual zoning designation from Black Knight
2. Have a dual zoning designation from Black Knight but one of the zoning codes does not match the results from the spatial join.

First, isolate these observations and separate each zoning code into its own observation. 
**NOTE:** For Arlington, the max codes was 4- adjust the `into = ` argument accordingly for each jurisdiction.

```{r real-final}

dual <- jurtotal %>% 
  filter((str_detect(zoning, "/") == TRUE & 
            str_detect(zoning_bk, ZN_DESIG) == FALSE) |
          (str_detect(zoning, "/") == TRUE &
           missing_coord == TRUE)) %>% 
  separate(zoning_bk, 
           into = c("z1", "z2", "z3", "z4"),
           sep = "/",
           remove = FALSE)

```
Get rid of values that are NOT in the cleaned zoning code data. These codes do not allow for residential development.

```{r no-non-res}

dual <- dual %>% 
  mutate_at(vars(z1, z2, z3, z4),
            ~ ifelse(!. %in% zoningcode1$zoning_code,
                     NA,
                     .))

```

Convert to long so that we can list the zoning variables- then `group_by` the parcel ID, and check the lengths to make sure that is unique, so that we can merge results back in.

```{r long}

long <- dual %>% 
  gather(key = "position", value = "code", -c(1:47, 52))

dual2 <- long %>% 
  group_by(assessorsparcelnumberapnpin) %>% 
  summarize(zlist = list(code))

stopifnot(nrow(dual) == nrow(dual2))

```

If there is only one code left after getting rid of the ones that don't allow for residential codes, take that code.

```{r true1}
dual2 <- dual2 %>% 
  mutate(count = map(zlist, ~ sum(!is.na(.)))) %>% 
  mutate(zoning_true = ifelse(count == 1,
                              map(zlist, max, na.rm = TRUE),
                              ""))
```

Priotize codes that are purely residential (begin with R). If there are more than 1 "R" zoning codes, we want to take the one with the lower number (take the lower density code, so we do not overestimate, since part of the building will be in the lower density area). We do this by extracting the numbers from the zoning codes, taking the minimum, and then selecting the element that contains that number.

```{r true2, warning=FALSE}

dual2 <- dual2 %>% 
  mutate(res = map(zlist, ~ substr(., 1, 1)),
         reslist = map(zlist, ~ .[str_detect(., "R")]),
         num1 = map(zlist, parse_number),
         num2 = map(num1, abs),
         num3 = map(num2, ~ round(min(., na.rm = TRUE), 0)),
         num4 = as.character(num3),
         r_count = map(res, ~ sum(. == "R", na.rm = TRUE)),
         zoning_true = ifelse(r_count == 1 & zoning_true == "",
                              map(zlist, ~ max(.[str_detect(., "R")],
                                               na.rm = TRUE)),
                              zoning_true),
         zoning_true = ifelse(r_count == 2,
                              map2(reslist, num4, ~ max(.[str_detect(.x, c(.y, "R"))],
                                                      na.rm = TRUE)),
                              zoning_true))

```

The remaining codes only allow for one dwelling districts, so we will just take the first code.

```{r true-last}

dual3 <- dual2 %>% 
  mutate(zoning_true = ifelse(zoning_true == "",
                       map(zlist, 1),
                       zoning_true),
         zoning_true = as.character(zoning_true))

stopifnot(sum(is.na(dual3$zoning_true)) == 0)
```

Look at a few observations to confirm.

```{r sample-weird}

dual3 %>% 
  sample_n(10) %>% 
  select(parcel = assessorsparcelnumberapnpin, zlist, zoning_true) %>% 
  knitr::kable()

```

Select only the clean zoning variable, merge back into main dataset, and replace zoning variable. Initial Black Knight variable will still be archived.

```{r recomb}

jurfinal <-jurtotal %>% 
  left_join(select(dual3,
                   assessorsparcelnumberapnpin, zoning = zoning_true),
            by = "assessorsparcelnumberapnpin") %>% 
  rename(zoning= zoning.x) %>% 
  mutate(zoning = ifelse(!is.na(zoning.y),
                         zoning.y,
                         zoning)) %>% 
  select(-zoning.y)

```

Save the final parcel data so it can be used for the special exceptions (this is only required for Arlington county).

```{r save-vacant}

saveRDS(jurfinal, "L:/Libraries/RegHsg/Data/arlington/arlington-lots.Rdata")
saveRDS(zoningcode1, "L:/Libraries/RegHsg/Data/arlington/arlington-clean-zoning-code.Rdata")

```


** Switch to `04_vacant-lots-2-arlington-se.Rmd` **


Clear global environment.

```{r}

rm(list=ls())

```
